---
title: "Figure 5 and supplementary figure 5"
author: "Niccolò Arecco, Ivano Mocavini and Enrique Blanco"
toc: true
number-sections: true
highlight-style: pygments
format:
  html:
    embed-resources: true
    code-fold: true
    code-copy: true
    theme:
      light: flatly
      dark: darkly
editor: source
bibliography: ../Paper_Analysis_References.bib
---

Last code execution: `r format(Sys.time(), '%Y %B %d, %A @ %H:%m:%S')`.

# Intro

ChIP-seq analysis.

```{r, empty, eval = FALSE, echo = FALSE}
# this is just to have r chunks inserted with 
```

To fetch some of the inclusion tables (PSI data) stored on the CRG cluster I mount the server on my local computer with the following command in the terminal.

```{bash, crgmnt, eval = F, echo = T}
sshfs narecco@ant-login.linux.crg.es:/users/mirimia ~/mnt
```

which prompts my password and then allows me to navigate the cluster directories.

# Set Up

General info.

## Packages

Load packages required for the analysis and suppress any message. Check the @sec-sessioninfo section at the end for package versions.

```{r, Loading required packages, warning = T, message = FALSE}
library(dplyr, warn.conflicts = F, quietly = T)
library(readr)
library(readxl)
library(stringr)
library(tidyr)
library(tibble)
library(ggplot2)
library(ggforce)
library(ggsignif)
library(ggh4x)
library(ggrastr) # for rasterizing only the geom_point layer in a plot with many data-points
library(rstatix)
library(janitor)
library(DT)
library(BioVenn) # Venn Diagram
library(caroline) # Spie chart
library(patchwork)
```

## Functions

Define a plot style.
```{r, label = "ggplot themes", eval = T}
theme_classic(base_family = 'Arial', base_size = 6) +
  theme(axis.text = element_text(colour = 'black'),
        axis.ticks.length = unit(1, 'mm'),
        axis.ticks = element_line(linewidth = 0.1),
        axis.line = element_line(linewidth = 0.1),
        axis.title = element_text(size = 5),
        panel.grid.major = element_line(linewidth = 0.2, colour = 'gray84'),
        strip.text = element_text(size = 10),
        strip.background = element_blank(),
        legend.title = element_blank(), 
        legend.key.size = unit(1, 'mm'),
        legend.position = c(0.21, 0.92),
        legend.box.background = element_blank(),
        legend.margin = margin(t = -1, b = 1, unit = 'mm'),
        plot.background = element_blank(),
        panel.background = element_blank()) -> scatter_th

theme_classic(base_family = 'Arial', base_size = 6) +
  theme(axis.text = element_text(colour = 'black'),
        axis.text.x = element_text(angle = 45, hjust = 1, margin = margin(t = -0.4, unit = 'mm') ),
        axis.title.x = element_blank(),
        axis.ticks.length.y = unit(1, 'mm'),
        axis.ticks = element_line(linewidth = 0.1),
        axis.ticks.x = element_blank(),
        axis.line = element_line(linewidth = 0.1),
        axis.title = element_text(size = 5),
        panel.grid.major.y = element_line(linewidth = 0.2, colour = 'gray84'),
        panel.background = element_blank(),
        panel.spacing = unit(3, 'mm'),
        strip.text = element_text(size = 5),
        strip.background = element_rect(linewidth = 0.1),
        plot.background = element_blank() ) -> boxplot_th

theme_classic(base_family = 'Arial', base_size = 6) +
  theme(axis.text = element_text(colour = 'black'),
        axis.text.x = element_text(margin = margin(t = -0.3, unit = 'mm') ),
        axis.ticks.length.y = unit(1, 'mm'),
        axis.ticks = element_line(linewidth = 0.1),
        axis.ticks.x = element_blank(),
        axis.line = element_line(linewidth = 0.1),
        axis.title = element_text(size = 5),
        panel.grid.major = element_line(linewidth = 0.2, colour = 'gray84'),
        panel.background = element_blank(),
        panel.spacing = unit(3, 'mm'),
        legend.position = c(0.93, 0.87),
        legend.key.size = unit(1, 'mm'),
        legend.title = element_blank(),
        legend.box.background = element_blank(),
        legend.margin = margin(t = -1, b = 1, unit = 'mm'),
        strip.text = element_text(size = 5),
        strip.background = element_rect(linewidth = 0.1),
        plot.background = element_blank() ) -> profile_th
```


Function import GENEprofile files from SeqCode

```{r, fun import GENEprofile files from SeqCode}
# read "GENEprofile" file types from SeqCode
read_prfl <- function(path_dir, txt_name, epitope, geneset_type, sample_name) {
  df_in <- read_delim(file = file.path(path_dir, txt_name), 
                      delim = '\t', show_col_types = FALSE, col_names = F ) |>
    setNames( c("MetaPosition", paste0(epitope, "_signal") ) ) |>
    mutate(type = geneset_type, Sample = sample_name ) 
  return(df_in)
}

```

## Directories & File Paths

Here I organise all the file and directories- paths I need to run the analysis and define where to save the processed tables and figures.

```{r, set paths and dicretories, warning = FALSE, message = FALSE}
oneDrive_Dir <- file.path("~/OneDrive - CRG - Centre de Regulacio Genomica/Suz12_AS_project")  
code_dir_fig <- file.path(oneDrive_Dir, "_Code/Fig5")
tbl_dir_fig5 <- file.path(code_dir_fig, "tables")
pdf_dir_fig5 <- file.path(code_dir_fig, "pdfs")

if (!dir.exists(pdf_dir_fig5)) { dir.create(pdf_dir_fig5, recursive = T) }
if (!dir.exists(tbl_dir_fig5)) { dir.create(tbl_dir_fig5, recursive = T) }
```

Path to DropBox folders where Enrique releases the documents.

```{r, paths to dropbox}
dropbox_dir <- file.path('~/Dropbox (CRG ADV)/54_Suz12AS')

round2_h3k27me3_peaks_dir <- file.path(dropbox_dir, 'ROUND2_H3K27me3_ChIP/GENESETS/MACS1')
stopifnot(dir.exists(round2_h3k27me3_peaks_dir))
round3_suz12_peaks_dir <- file.path(dropbox_dir, 'ROUND3 - SUZ12 ChIP/files/genesets')
stopifnot(dir.exists(round3_suz12_peaks_dir))

round17_meta_dir <- file.path(dropbox_dir, 'ROUND17_Figure3/01_files_for_boxplotofmetaplots')
round17_scatter_dir <- file.path(dropbox_dir, 'ROUND17_Figure3/02_files_for_scattersboxplots')
stopifnot(dir.exists(round17_meta_dir))
stopifnot(dir.exists(round17_scatter_dir))

round21_meta_dir <- file.path(dropbox_dir, 'ROUND21_PcGSubUnitChIPseq/METAPLOTS/GENEPLOTS')
stopifnot(dir.exists(round21_meta_dir))
round23_scatter_dir <- file.path(dropbox_dir, 'ROUND23_FinalFiguresRevision/SCATTERPLOTS')
stopifnot(dir.exists(round23_scatter_dir))
round23_metagene_dir <- file.path(dropbox_dir, 'ROUND23_FinalFiguresRevision/METAPLOTS')
stopifnot(dir.exists(round23_metagene_dir))
```

List all bed files to be used for the metagene profiles
```{r, bed files paths}
bed_paths <- list.files(round17_meta_dir, recursive = T, pattern = "*.bed$")
```

Path where the data to plot meta genes profiles for H3K27me3 are stored.
```{r, Processed data metaplots CGI dir path}
CGI_Dir <- file.path(oneDrive_Dir, '11_ChIP/Processed_data/Metaplots_CGI')
stopifnot(dir.exists(CGI_Dir))
```

# Main Figure Panels

## Spike-in normalised H3K27me3 signal in 2.5kb genomic bins in endogenous cell lines

Import all 2.5 kb bins of H3K27me3.

```{r, import H3K27me3 bins of 2.5kb bins endogenous cell lines}
all_bins25kb_K27me3 <- read_delim(file = file.path(round17_scatter_dir, 'H3K27me3_R3all_bins_2.5Kb_yes.txt'),
                                  delim = '\t', progress = F,
                                  col_names = c('bin', "WT", 'CSex4', 'dex4', 'KO'),
                                  show_col_types = FALSE) |>
  mutate(chr = str_split_fixed(pattern = "\\*", string = bin, n = 3)[,1],
         start = str_split_fixed(pattern = "\\*", string = bin, n = 3)[,2],
         end = str_split_fixed(pattern = "\\*", string = bin, n = 3)[,3]) |>
  relocate(chr, start, end, .after = bin) 
```

Import the annotated bins that belong to SUZ12 targeted CGI and those that contain a CGI that is not target

```{r, import bins of CGIs and others}
bins25kbK27me3_CGI_SUZ12 <- read_delim(
  file = file.path(round17_scatter_dir, 'H3K27me3_R3all_bins_2.5Kb_yes_CGI-SUZ12.txt'),
  delim = c(' '), col_names = c('bin', "garbage"), show_col_types = FALSE
  )

bins25kbK27me3_other_CGI <- read_delim(
  file = file.path(round17_scatter_dir, 'H3K27me3_R3all_bins_2.5Kb_yes_CGI-NOPCG.txt'), 
  delim = c(' '), col_names = c('bin', "garbage"), show_col_types = FALSE
  )
```

Check how many of the total bins are in each set (SUZ12 targeted CGIs or others CGIs). 
```{r, check imported bins}
table(all_bins25kb_K27me3$bin %in% bins25kbK27me3_CGI_SUZ12$bin)
table(all_bins25kb_K27me3$bin %in% bins25kbK27me3_other_CGI$bin)
```

Map the total bins in each set
```{r, map bins}
all_bins25kb_K27me3 |>
  mutate(type = case_when(bin %in% bins25kbK27me3_CGI_SUZ12$bin ~ "PcG targeted CGI",
                          bin %in% bins25kbK27me3_other_CGI$bin ~ "Other CGI") ) |>
  mutate(type = ifelse(is.na(type), yes = "Rest of the genome", no = type )
         ) -> all_bins25kb_K27me3
```

Set bin type as factor.
```{r, factorise bins types}
all_bins25kb_K27me3$type <- factor(all_bins25kb_K27me3$type,
                                   levels = c("PcG targeted CGI", "Other CGI", "Rest of the genome") )
```

Overview of bins file.
```{r, data table K36}
datatable(head(all_bins25kb_K27me3, 150), rownames = FALSE, filter = 'top', 
          options = list(pageLength = 5, autoWidth = TRUE) )
```

Define the flanking bins (-1 and +1 bins) surrounding CGI.
```{r, PcG target index}
indx_targeted <- which(all_bins25kb_K27me3$type == "PcG targeted CGI")
message(length(indx_targeted), ' PcG targeted bins')
```

Get first the +1 bins:

If the difference in index number between the targeted bin and the + 1 is 1, meaning the 2 PcG targeted CGI are one next to each other, exclude the +1 bins. Otherwise, label every +1 index number as downstream flanking PcG targeted CGI.

```{r, get flanking downstream bins}
plus_one_targeted <- c()
for (i in 1:(length(indx_targeted)-1) ) {
  
  if ( (indx_targeted[i + 1] - indx_targeted[i]) == 1 ) {
    # Do no nothing
    plus_one_targeted[i] <- 0
    
  } else {
    plus_one_targeted[i] <- indx_targeted[i] + 1
  }
}
```

Here's an example 
```{r, example of plus one selection}
message('Targeted indexes:\t', paste0(head(indx_targeted, 7), sep = ' ') )
message('Flanking +1 indexes:\t', paste0(head(plus_one_targeted, 7), sep = ' ') )
```

Since bins numbered `2647` and `2648` are next to each other, the bin numbered `2649` is selected and there's a `0` for the `2648`. By removing the zeros I remove the consecutive CGI bins.

```{r, remove plus one consecutive bins}
plus_one_targeted <- plus_one_targeted[plus_one_targeted != 0]
```

Get the -1 bins:
here the logic is the same as above, but inverted. If the difference in index number between the targeted bin and the -1 bin is 1, meaning the 2 PcG targeted CGI are one next to each other, exclude the -1 bins.
```{r, get flanking upstream bins}
minus_one_targeted <- c()
for (i in 2:(length(indx_targeted)) ) {
  
  if ( ( indx_targeted[i] - indx_targeted[i - 1] ) == 1 ) {
    # Do no nothing
    minus_one_targeted[i] <- 0
    
  } else {
    minus_one_targeted[i] <- indx_targeted[i] - 1
  }
}
```

Remove first NA, and consecutive bins defined by zeros.
```{r, remove NA}
minus_one_targeted[which(  is.na(minus_one_targeted) )] <- indx_targeted[1] - 1
minus_one_targeted <- minus_one_targeted[minus_one_targeted != 0]
```

Check binning
```{r, example of minus one selection}
message('Targeted indexes:\t', paste0(head(indx_targeted, 7), sep = ' ') )
message('Flanking -1 indexes:\t', paste0(head(minus_one_targeted, 7), sep = ' ') )
```

Since bins numbered `2647` and `2648` are next to each other, the bin numbered `2646` is selected and there's a `0` for the `2647`. By removing the zeros I remove the consecutive CGI bins.

Check that there's no overlap between the targeted and the flanking regions.
```{r, check overalp of plus or minus with tarted}
table(plus_one_targeted %in% indx_targeted)
table(minus_one_targeted %in% indx_targeted)
```

Check that some flanking bins can be both upstream or downstream.

```{r, check overalp between plus and minus}
table(minus_one_targeted %in% plus_one_targeted)
```

Create a new bin type variable called `type2`.
```{r, assign new GI type}
all_bins25kb_K27me3$type2  <- NA
all_bins25kb_K27me3[ indx_targeted, ]$type2 <- "PcG targeted CGI"
all_bins25kb_K27me3[ plus_one_targeted, ]$type2 <- "PcG targeted CGI + 1"
all_bins25kb_K27me3[ minus_one_targeted, ]$type2 <- "PcG targeted CGI - 1"
all_bins25kb_K27me3[ is.na(all_bins25kb_K27me3$type2), ]$type2 <- "Rest of the genome"  

all_bins25kb_K27me3$type2 <- factor(all_bins25kb_K27me3$type2,
                                   levels = c(
                                     "PcG targeted CGI", 
                                     "PcG targeted CGI + 1",
                                     "PcG targeted CGI - 1",
                                     "Rest of the genome") )
table(all_bins25kb_K27me3$type2)
```

Log10 transform the bins signal.
```{r, Log10 transform the bins signal}
all_bins25kb_K27me3 |>
  column_to_rownames(var = "bin") |>
  select(chr, start, end, WT, CSex4, dex4, KO, type, type2) |>
  mutate( across( where(is.double), log10 ) ) |>
  # arrange is used to favour the CGI point to appear later in the plot and therefore be above the other points.
  arrange( desc(type2) ) -> log_bins25kb_K27me3
```

Check total number of bin.
```{r, bin num endogenous}
num_all_bins_2.5kb <- nrow(log_bins25kb_K27me3)
message("Number of bins of 2.5kb width to plot: ",  num_all_bins_2.5kb)
```

Individual bin types numbers:
```{r, check number of bins in each set endogenous lines}
targeted_nBins <- table(log_bins25kb_K27me3$type2)[names(table(log_bins25kb_K27me3$type2))=="PcG targeted CGI"]
plus_nBins <- table(log_bins25kb_K27me3$type2)[names(table(log_bins25kb_K27me3$type2))=="PcG targeted CGI + 1"]
minus_nBins <- table(log_bins25kb_K27me3$type2)[names(table(log_bins25kb_K27me3$type2))=="PcG targeted CGI - 1"]
flanking_nBins <- sum(plus_nBins, minus_nBins)
ROTG_nBins <- table(log_bins25kb_K27me3$type2)[names(table(log_bins25kb_K27me3$type2))=="Rest of the genome"]

message("Bins containing a targeted CGI: ", targeted_nBins, "\n",
        "Bins flanking a targeted CGI: ", flanking_nBins, "\n",
        "Bins in the rest of genome: ", ROTG_nBins)
```

### Scatter plot ∆ex4 vs WT

Fit a linear model between different conditions.

```{r, linear fit WT vs dex4}
intercet_wt_dex4 <- coef( lm(log_bins25kb_K27me3$dex4 ~ log_bins25kb_K27me3$WT ) )[1] 
slope_wt_dex4 <- coef( lm(log_bins25kb_K27me3$dex4 ~ log_bins25kb_K27me3$WT ) )[2] 
```

If I were to plot all bins signal as a scatter plot of WT vs ∆ex4 I would do something like this:
```{r, plot all bins wt vs dex4, eval = FALSE, echo = TRUE}
log_bins25kb_K27me3 |>
  ggplot() +
  aes(x = WT, y = dex4, colour = type2) +
  # facet_wrap(~type2, scales = 'fixed') +
  geom_point(size = 0.75) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_dex4, intercept = intercet_wt_dex4, linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5) ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "PcG targeted CGI" = '#A2234C',
                                 "Rest of the genome" = '#72817F') ) +
  labs(x = expression("Spike-in normalized WT H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized ∆ex4 H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  scatter_th -> pBins_WT_vs_dex4_All_Targeted

# do not plot cause it's 2 many points
# pBins_WT_vs_dex4_All_Targeted
ggsave(path = pdf_dir_fig5,
       filename = paste0("Fig5B_CGI_coloured_WT_vs_Dex4_bins_n", num_all_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_dex4_All_Targeted, device = cairo_pdf, units = "cm",
       width = 4.5, height = 4.5)
```

However that breaks the RStudio graphic engine.

To overcome some of the overplotting I remove points that are identical by `type2` value. If 2 point have the exact same values, the duplicated point is removed from the dataframe. This helps removing points that are going to be plotted on top of each other, making the plot more manageable. This operation requires the use of the dataframe row names.

Start by removing the many "rest of the genome" bins
```{r, filter rest of genome}
indx_dups_ROTG_wt_dex4 <-
  which(duplicated(log_bins25kb_K27me3[log_bins25kb_K27me3$type2 == "Rest of the genome"  , c("WT", "dex4")]))
row_names_dups_ROTG <- rownames(log_bins25kb_K27me3[indx_dups_ROTG_wt_dex4, ]) 
wt_dex4_bins <- log_bins25kb_K27me3[!rownames(log_bins25kb_K27me3) %in% row_names_dups_ROTG,]
```

Then removing the "PcG targeted CGI"
```{r, filter CGI}
indx_dups_PTC_wt_dex4 <- which(duplicated(wt_dex4_bins[wt_dex4_bins$type2 == "PcG targeted CGI", c("WT", "dex4")] ) )
row_names_dups_PTC <- rownames(wt_dex4_bins[indx_dups_PTC_wt_dex4, ]) 
wt_dex4_bins <- wt_dex4_bins[!rownames(wt_dex4_bins) %in% row_names_dups_PTC,]
```

Lastly the +1 and -1 bins

```{r, filter flanking bins}
indx_dups_P1M1_wt_dex4 <- which(duplicated(wt_dex4_bins[wt_dex4_bins$type2 %in% c("PcG targeted CGI + 1", "PcG targeted CGI + 1"), c("WT", "dex4")] ) )
row_names_dups_P1M1 <- rownames(wt_dex4_bins[indx_dups_P1M1_wt_dex4, ]) 
wt_dex4_bins <- wt_dex4_bins[!rownames(wt_dex4_bins) %in% row_names_dups_P1M1,]
```

Re-calculate bin numbers now after filtering

```{r, re-get number of bin}
num_fltrd_bins_2.5kb <- nrow(wt_dex4_bins)
message("Number of bins of 2.5kb width to plot: ",  num_fltrd_bins_2.5kb)
```

Plot figure 5C.

```{r, plot filtered scatter plot dex4 fig 5C}
ggplot(wt_dex4_bins) +
  aes(x = WT, y = dex4, colour = type2) +
  # facet_wrap(~type2, scales = 'fixed') +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_dex4, intercept = intercet_wt_dex4, linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                  "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized ∆ex4 H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th -> pBins_WT_vs_dex4_Filtered_Targeted
```

Save to pdf.

```{r, save figure 5C as PDF}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5C_CGI_coloured_WT_vs_Dex4_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_dex4_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

::: {.callout-note}
The scatter plot contains too many points to be efficiently plotted. For an effective scatter plot visualisation I rasterise only the specific points layer of the `ggplot2` plot and keep all other labels and text in vector format.
:::

```{r, rasterise dex4}
rBins_WT_vs_dex4_Filtered_Targeted <- rasterize(pBins_WT_vs_dex4_Filtered_Targeted, layers = 'Point', dpi = 400)
```

Plot rasterised scatter plot for figure 5C
```{r, display figure 5C as rasterised, fig.width = 2, fig.height = 2}
rBins_WT_vs_dex4_Filtered_Targeted
```

Save to pdf.

```{r, save raterised dex4 fig 5C}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5C_RASTERIZED_CGI_coloured_WT_vs_Dex4_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_dex4_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Now do the same but for CSex4 vs WT.

### Scatter plot CSex4 vs WT

Get linear fit.
```{r, linear fit WT vs CSex4}
intercet_wt_CSex4 <- coef( lm(log_bins25kb_K27me3$CSex4 ~ log_bins25kb_K27me3$WT ) )[1] 
slope_wt_CSex4 <- coef( lm(log_bins25kb_K27me3$CSex4 ~ log_bins25kb_K27me3$WT ) )[2] 
```

Filter the bins like before.
Start by removing the many "rest of the genome" bins

```{r, filter CSex4 RotG }
indx_dups_ROTG_wt_CSex4 <- which(duplicated(log_bins25kb_K27me3[log_bins25kb_K27me3$type2 == "Rest of the genome", c("WT", "CSex4")]))
row_names_dups_ROTG <- rownames(log_bins25kb_K27me3[indx_dups_ROTG_wt_CSex4, ]) 
wt_CSex4_bins <- log_bins25kb_K27me3[!rownames(log_bins25kb_K27me3) %in% row_names_dups_ROTG,]
```

Filter the "PcG targeted CGI" in CSex4 vs WT dataset
```{r, filter CSex4 CGI}
indx_dups_PTC_wt_CSex4 <- which(duplicated(wt_CSex4_bins[wt_CSex4_bins$type2 == "PcG targeted CGI", c("WT", "CSex4")] ) )
row_names_dups_PTC <- rownames(wt_CSex4_bins[indx_dups_PTC_wt_CSex4, ]) 
wt_CSex4_bins <- wt_CSex4_bins[!rownames(wt_CSex4_bins) %in% row_names_dups_PTC,]
```

And the +1 and -1 flanking bins
```{r, filter CSex4 flanking bins}
indx_dups_P1M1_wt_CSex4 <- which(duplicated(wt_CSex4_bins[wt_CSex4_bins$type2 %in% c("PcG targeted CGI + 1", "PcG targeted CGI + 1"), c("WT", "CSex4")] ) )
row_names_dups_P1M1 <- rownames(wt_CSex4_bins[indx_dups_P1M1_wt_CSex4, ]) 
wt_CSex4_bins <- wt_CSex4_bins[!rownames(wt_CSex4_bins) %in% row_names_dups_P1M1,]
```

Check number of bins after filtering
```{r, get CSex4 bins}
num_fltrd_bins_2.5kb <- nrow(wt_CSex4_bins)
message("Number of bins of 2.5kb width to plot: ", num_fltrd_bins_2.5kb)
```

Basically this filtering removes duplicated points.
```{r, plot filtered scatter plot CSex4 vs WT}
ggplot(wt_CSex4_bins) +
  aes(x = WT, y = CSex4, colour = type2) +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_CSex4, intercept = intercet_wt_CSex4, 
              linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                 "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized CSex4 H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th -> pBins_WT_vs_CSex4_Filtered_Targeted
```

Save to pdf vector.

```{r, save CSex4 vs WT scatterplot vector}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("Fig5B_CGI_coloured_WT_vs_CSex4_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_CSex4_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise only the point layer of ggplot

```{r, rasterise CSex4 vs WT scatterplot}
rBins_WT_vs_CSex4_Filtered_Targeted <- rasterize(pBins_WT_vs_CSex4_Filtered_Targeted, layers = 'Point', dpi = 400)
```

Show plot as presented in figure 5B.

```{r, display rasterised figure 5B, fig.width = 2, fig.height = 2}
rBins_WT_vs_CSex4_Filtered_Targeted
```

Save to pdf rasterised scatter plot.

```{r, save CSex4 vs WT scatterplot rasterised}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("Fig5B_RASTERIZED_CGI_coloured_WT_vs_CSex4_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_CSex4_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

## Spike-in normalised H3K27me3 signal in 2.5kb genomic bins in KO rescues cell lines

Here I repeat the same analysis but for the SUZ12-KO isoform specific rescue cell lines.

```{r, import rescues}
all_bins25kb_K27me3_rescues <- read_delim(
  file = file.path(round23_scatter_dir, 'H3K27me3_R3all--W2-S1-S3-L4-L6-L3D2-S3D2_bins_2.5Kb_yes.txt'),
                                  delim = ' ', progress = F,
                                  col_names = c('bin', 'WT', 'CSex4', 'dex4', 'KO',
                                                'WT2', 'KOrS1', 'KOrS3', 'KOrL4', 'KOrL6', 'KOrL3D2', 'KOrS3D2'),
                                  show_col_types = FALSE) |>
  mutate(chr = str_split_fixed(pattern = "\\*", string = bin, n = 3)[,1],
         start = str_split_fixed(pattern = "\\*", string = bin, n = 3)[,2],
         end = str_split_fixed(pattern = "\\*", string = bin, n = 3)[,3]) |>
  relocate(chr, start, end, .after = bin) 
```

The samples `'WT2', 'KOrS1', 'KOrS3', 'KOrL4', 'KOrL6', 'KOrL3D2', 'KOrS3D2'` were done after in a second round compared to the first round of samples (`'WT', 'CSex4', 'dex4', 'KO'`). 

Check how many bins fall within a SUZ12 CGIs or other CGI

```{r}
table(all_bins25kb_K27me3_rescues$bin %in% bins25kbK27me3_CGI_SUZ12$bin)
table(all_bins25kb_K27me3_rescues$bin %in% bins25kbK27me3_other_CGI$bin)
```

Annotate rescues bins that fall in targeted CGIs.

```{r}
all_bins25kb_K27me3_rescues |>
  mutate(type = case_when(bin %in% bins25kbK27me3_CGI_SUZ12$bin ~ "PcG targeted CGI",
                          bin %in% bins25kbK27me3_other_CGI$bin ~ "Other CGI") ) |>
  mutate(type = ifelse(is.na(type), yes = "Rest of the genome", no = type )
  ) -> all_bins25kb_K27me3_rescues
```

```{r}
all_bins25kb_K27me3_rescues$type <- factor(all_bins25kb_K27me3_rescues$type,
                                   levels = c("PcG targeted CGI", "Other CGI", "Rest of the genome") )
```

Same as before identify the CGI flanking bins.

First the downstream +1 bins.

```{r}
indx_targeted <- which(all_bins25kb_K27me3_rescues$type == "PcG targeted CGI")

plus_one_targeted <- c()
for (i in 1:(length(indx_targeted)-1) ) {
  
  if ( (indx_targeted[i + 1] - indx_targeted[i]) == 1 ) {
    # Do no nothing
    plus_one_targeted[i] <- 0
    
  } else {
    plus_one_targeted[i] <- indx_targeted[i] + 1
  }
}

plus_one_targeted <- plus_one_targeted[plus_one_targeted != 0]
```

Then the upstream -1 bins

```{r}
minus_one_targeted <- c()
for (i in 2:(length(indx_targeted)) ) {
  
  if ( ( indx_targeted[i] - indx_targeted[i - 1] ) == 1 ) {
    # Do no nothing
    minus_one_targeted[i] <- 0
    
  } else {
    minus_one_targeted[i] <- indx_targeted[i] - 1
  }
}

minus_one_targeted[which(  is.na(minus_one_targeted) )] <- indx_targeted[1] - 1
minus_one_targeted <- minus_one_targeted[minus_one_targeted != 0]
```

Create a new bin type variable called `type2` also for rescue data.

```{r, assign new CGI type in rescues}
all_bins25kb_K27me3_rescues$type2  <- NA
all_bins25kb_K27me3_rescues[ indx_targeted, ]$type2 <- "PcG targeted CGI"
all_bins25kb_K27me3_rescues[ plus_one_targeted, ]$type2 <- "PcG targeted CGI + 1"
all_bins25kb_K27me3_rescues[ minus_one_targeted, ]$type2 <- "PcG targeted CGI - 1"
all_bins25kb_K27me3_rescues[ is.na(all_bins25kb_K27me3_rescues$type2), ]$type2 <- "Rest of the genome"  

all_bins25kb_K27me3_rescues$type2 <- factor(all_bins25kb_K27me3_rescues$type2,
                                    levels = c(
                                      "PcG targeted CGI", 
                                      "PcG targeted CGI + 1",
                                      "PcG targeted CGI - 1",
                                      "Rest of the genome") )
table(all_bins25kb_K27me3_rescues$type2)
```

Log10 transform the bins signal in rescue data.
```{r, Log10 transform the bins signal in rescues}
all_bins25kb_K27me3_rescues |>
  column_to_rownames(var = "bin") |>
  select(chr, start, end, WT2, KOrS1, KOrS3, KOrL4, KOrL6, KOrL3D2, KOrS3D2, type, type2) |>
  mutate( across( where(is.double), log10 ) ) |>
  # arrange is used to favour the CGI point to appear later in the plot and therefore be above the other points.
  arrange( desc(type2) ) -> log_bins25kb_K27me3_rescue
```

Check total number of bin in rescues.
```{r, bin num rescues}
num_all_bins_2.5kb <- nrow(log_bins25kb_K27me3_rescue)
message("Number of bins of 2.5kb width to plot: ",  num_all_bins_2.5kb)
```

Individual bin types numbers in rescues:
```{r, check number of bins in each set rescues}
targeted_nBins <- table(log_bins25kb_K27me3_rescue$type2)[names(table(log_bins25kb_K27me3_rescue$type2))=="PcG targeted CGI"]
plus_nBins <- table(log_bins25kb_K27me3_rescue$type2)[names(table(log_bins25kb_K27me3_rescue$type2))=="PcG targeted CGI + 1"]
minus_nBins <- table(log_bins25kb_K27me3_rescue$type2)[names(table(log_bins25kb_K27me3_rescue$type2))=="PcG targeted CGI - 1"]
flanking_nBins <- sum(plus_nBins, minus_nBins)
ROTG_nBins <- table(log_bins25kb_K27me3_rescue$type2)[names(table(log_bins25kb_K27me3_rescue$type2))=="Rest of the genome"]

message("Bins containing a targeted CGI: ", targeted_nBins, "\n",
        "Bins flanking a targeted CGI: ", flanking_nBins, "\n",
        "Bins in the rest of genome: ", ROTG_nBins)
```

### Scatter plot KO+SUZ12-S vs WT

Fit a linear model between different conditions.

```{r, linear fit WT vs KOrS}
intercet_wt_KOrS1 <- coef( lm(log_bins25kb_K27me3_rescue$KOrS1 ~ log_bins25kb_K27me3_rescue$WT2 ) )[1] 
slope_wt_KOrS1 <- coef( lm(log_bins25kb_K27me3_rescue$KOrS1 ~ log_bins25kb_K27me3_rescue$WT2 ) )[2] 

intercet_wt_KOrS3 <- coef( lm(log_bins25kb_K27me3_rescue$KOrS3 ~ log_bins25kb_K27me3_rescue$WT2 ) )[1] 
slope_wt_KOrS3 <- coef( lm(log_bins25kb_K27me3_rescue$KOrS3 ~ log_bins25kb_K27me3_rescue$WT2 ) )[2] 
```

To overcome some of the overplotting I remove points that are identical by `type2` value. If 2 point have the exact same values, the duplicated point is removed from the dataframe. 

Start by removing the many "rest of the genome" bins
```{r, filter rest of genome rescues KOrS1}
indx_dups_ROTG_wt_KOrS1 <-
  which(duplicated(log_bins25kb_K27me3_rescue[log_bins25kb_K27me3_rescue$type2 == "Rest of the genome"  , c("WT2", "KOrS1")]))
row_names_dups_ROTG <- rownames(log_bins25kb_K27me3_rescue[indx_dups_ROTG_wt_KOrS1, ]) 
wt_KOrS1_bins <- log_bins25kb_K27me3_rescue[!rownames(log_bins25kb_K27me3_rescue) %in% row_names_dups_ROTG,]
```

Then removing the "PcG targeted CGI"
```{r, filter CGI rescues KOrS1}
indx_dups_PTC_wt_KOrS1 <- which(duplicated(wt_KOrS1_bins[wt_KOrS1_bins$type2 == "PcG targeted CGI", c("WT2", "KOrS1")] ) )
row_names_dups_PTC <- rownames(wt_KOrS1_bins[indx_dups_PTC_wt_KOrS1, ]) 
wt_KOrS1_bins <- wt_KOrS1_bins[!rownames(wt_KOrS1_bins) %in% row_names_dups_PTC,]
```

Lastly the +1 and -1 bins

```{r, filter flanking bins KOrS1}
indx_dups_P1M1_wt_KOrS1 <- which(duplicated(wt_KOrS1_bins[wt_KOrS1_bins$type2 %in% c("PcG targeted CGI + 1", "PcG targeted CGI + 1"), c("WT2", "KOrS1")] ) )
row_names_dups_P1M1 <- rownames(wt_KOrS1_bins[indx_dups_P1M1_wt_KOrS1, ]) 
wt_KOrS1_bins <- wt_KOrS1_bins[!rownames(wt_KOrS1_bins) %in% row_names_dups_P1M1,]
```

Re-calculate bin numbers now after filtering

```{r, re-get number of bin rescues KOrS1}
num_fltrd_bins_2.5kb <- nrow(wt_KOrS1_bins)
message("Number of bins of 2.5kb width to plot: ",  num_fltrd_bins_2.5kb)
```

Plot figure 5E

```{r, plot filtered scatter plot dex4}
ggplot(wt_KOrS1_bins) +
  aes(x = WT2, y = KOrS1, colour = type2) +
  # facet_wrap(~type2, scales = 'fixed') +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_KOrS1, intercept = intercet_wt_KOrS1, linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                  "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized KO+S H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th -> pBins_WT_vs_KOrS1_Filtered_Targeted
pBins_WT_vs_KOrS1_Filtered_Targeted
```

Save to pdf.

```{r, save figure 5D as PDF}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5E_CGI_coloured_WT_vs_KOrS1_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_KOrS1_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise the points to 400dpi

```{r, rasterise KOrS1}
rBins_WT_vs_KOrS1_Filtered_Targeted <- rasterize(pBins_WT_vs_KOrS1_Filtered_Targeted, layers = 'Point', dpi = 400)
```

Plot rasterised scatter plot for figure 3B
```{r, display figure 5E as rasterised, fig.width = 2, fig.height = 2}
rBins_WT_vs_KOrS1_Filtered_Targeted
```

Save to pdf.

```{r, save raterised 5E}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5E_RASTERIZED_CGI_coloured_WT_vs_KOrS1_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_KOrS1_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Now for the KO+L (KOrL4).

### Scatter plot KO+SUZ12-L vs WT

Get linear fit.
```{r, linear fit WT vs KOrL4 or KOrL6 }
intercet_wt_KOrL4 <- coef( lm(log_bins25kb_K27me3_rescue$KOrL4 ~ log_bins25kb_K27me3_rescue$WT2 ) )[1] 
slope_wt_KOrL4 <- coef( lm(log_bins25kb_K27me3_rescue$KOrL4 ~ log_bins25kb_K27me3_rescue$WT2 ) )[2] 

intercet_wt_KOrL6 <- coef( lm(log_bins25kb_K27me3_rescue$KOrL6 ~ log_bins25kb_K27me3_rescue$WT2 ) )[1] 
slope_wt_KOrL6 <- coef( lm(log_bins25kb_K27me3_rescue$KOrL6 ~ log_bins25kb_K27me3_rescue$WT2 ) )[2] 
```

Check linear fit parameters

```{r}
message("WT vs KO+L4: intercept = ", round(intercet_wt_KOrL4, 3), " slope = ", round(slope_wt_KOrL4, 3))
message("WT vs KO+L6: intercept = ", round(intercet_wt_KOrL6, 3), " slope = ", round(slope_wt_KOrL6, 3))
```

Filter the bins like before.
Start by removing the many "rest of the genome" bins

```{r, filter KOrL4 ROTG }
indx_dups_ROTG_wt_KOrL4 <- which(duplicated(log_bins25kb_K27me3_rescue[log_bins25kb_K27me3_rescue$type2 == "Rest of the genome", c("WT2", "KOrL4")]))
row_names_dups_ROTG <- rownames(log_bins25kb_K27me3_rescue[indx_dups_ROTG_wt_KOrL4, ]) 
wt_KOrL4_bins <- log_bins25kb_K27me3_rescue[!rownames(log_bins25kb_K27me3_rescue) %in% row_names_dups_ROTG,]
```

Filter the "PcG targeted CGI" in KO rescue Long clone 4 vs WT dataset

```{r, filter KOrL4 CGI}
indx_dups_PTC_wt_KOrL4 <- which(duplicated(wt_KOrL4_bins[wt_KOrL4_bins$type2 == "PcG targeted CGI", c("WT2", "KOrL4")] ) )
row_names_dups_PTC <- rownames(wt_KOrL4_bins[indx_dups_PTC_wt_KOrL4, ]) 
wt_KOrL4_bins <- wt_KOrL4_bins[!rownames(wt_KOrL4_bins) %in% row_names_dups_PTC,]
```

And the +1 and -1 flanking bins
```{r, filter KOrL4 flanking bins}
indx_dups_P1M1_wt_KOrL4 <- which(duplicated(wt_KOrL4_bins[wt_KOrL4_bins$type2 %in% c("PcG targeted CGI + 1", "PcG targeted CGI + 1"), c("WT2", "KOrL4")] ) )
row_names_dups_P1M1 <- rownames(wt_KOrL4_bins[indx_dups_P1M1_wt_KOrL4, ]) 
wt_KOrL4_bins <- wt_KOrL4_bins[!rownames(wt_KOrL4_bins) %in% row_names_dups_P1M1,]
```

Check number of bins after filtering:

```{r, get KOrL4 bins}
num_fltrd_bins_2.5kb <- nrow(wt_KOrL4_bins)
message("Number of bins of 2.5kb width to plot: ", num_fltrd_bins_2.5kb)
```

Plot figure 5D

```{r, plot filtered scatter plot KOrL4 vs WT}
ggplot(wt_KOrL4_bins) +
  aes(x = WT2, y = KOrL4, colour = type2) +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_KOrL4, intercept = intercet_wt_KOrL4, 
              linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                 "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized KO+L H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th -> pBins_WT_vs_KOrL4_Filtered_Targeted
pBins_WT_vs_KOrL4_Filtered_Targeted
```

Save to pdf vector.

```{r, save KOrL4 vs WT scatterplot vector}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("Fig5D_CGI_coloured_WT_vs_KOrL4_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_KOrL4_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise only the point layer of ggplot

```{r, rasterise KOrL4 vs WT scatterplot}
rBins_WT_vs_KOrL4_Filtered_Targeted <- rasterize(pBins_WT_vs_KOrL4_Filtered_Targeted, layers = 'Point', dpi = 400)
```

Show plot as presented in figure.

```{r, display rasterised figure 5E, fig.width = 2, fig.height = 2}
rBins_WT_vs_KOrL4_Filtered_Targeted
```

Save to pdf rasterised scatter plot.

```{r, save KOrL4 vs WT scatterplot rasterised}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("Fig5D_RASTERIZED_CGI_coloured_WT_vs_KOrL4_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_KOrL4_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

### Scatter plot KO+SUZ12-S3D vs WT

SUZ12-S 3D mutant H3K27me3 scatter plot.

Get linear fit.
```{r, linear fit WT2 vs KOrS3D2 }
intercet_wt_KOrS3D2 <- coef( lm(log_bins25kb_K27me3_rescue$KOrS3D2 ~ log_bins25kb_K27me3_rescue$WT2 ) )[1] 
slope_wt_KOrS3D2 <- coef( lm(log_bins25kb_K27me3_rescue$KOrS3D2 ~ log_bins25kb_K27me3_rescue$WT2 ) )[2] 
```

Check linear fit parameters

```{r}
message("WT2 vs KOrS3D2: intercept = ", round(intercet_wt_KOrS3D2, 3), 
        " slope = ", round(slope_wt_KOrS3D2, 3))
```

Filter the bins like before.
Start by removing the many "rest of the genome" bins

```{r, filter slope wt KOrS3D2 ROTG }
indx_dups_ROTG_wt_KOrS3D2 <- which(duplicated(log_bins25kb_K27me3_rescue[log_bins25kb_K27me3_rescue$type2 == "Rest of the genome", c("WT2", "KOrS3D2")]))
row_names_dups_ROTG <- rownames(log_bins25kb_K27me3_rescue[indx_dups_ROTG_wt_KOrS3D2, ]) 
wt_KOrS3D2_bins <- log_bins25kb_K27me3_rescue[!rownames(log_bins25kb_K27me3_rescue) %in% row_names_dups_ROTG,]
```

Filter the "PcG targeted CGI" in KO rescue SUZ12-Short 3D mutant vs WT dataset

```{r, filter KOrS3D2 CGI}
indx_dups_PTC_wt_KOrS3D2 <- which(duplicated(wt_KOrS3D2_bins[wt_KOrS3D2_bins$type2 == "PcG targeted CGI", c("WT2", "KOrS3D2")] ) )
row_names_dups_PTC <- rownames(wt_KOrS3D2_bins[indx_dups_PTC_wt_KOrS3D2, ]) 
wt_KOrS3D2_bins <- wt_KOrS3D2_bins[!rownames(wt_KOrS3D2_bins) %in% row_names_dups_PTC,]
```

And the +1 and -1 flanking bins
```{r, filter wt_KOrS3D2_bins flanking bins}
indx_dups_P1M1_wt_KOrS3D2 <- which(duplicated(wt_KOrS3D2_bins[wt_KOrS3D2_bins$type2 %in% c("PcG targeted CGI + 1", "PcG targeted CGI + 1"), c("WT2", "KOrS3D2")] ) )
row_names_dups_P1M1 <- rownames(wt_KOrS3D2_bins[indx_dups_P1M1_wt_KOrS3D2, ]) 
wt_KOrS3D2_bins <- wt_KOrS3D2_bins[!rownames(wt_KOrS3D2_bins) %in% row_names_dups_P1M1,]
```

Check number of bins after filtering:

```{r, get KOrS3D2 bins}
num_fltrd_bins_2.5kb <- nrow(wt_KOrS3D2_bins)
message("Number of bins of 2.5kb width to plot: ", num_fltrd_bins_2.5kb)
```

Plot figure S5G

```{r, plot filtered scatter plot KOrS3D2 vs WT, fig.width = 2, fig.height = 2}
ggplot(wt_KOrS3D2_bins) +
  aes(x = WT2, y = KOrS3D2, colour = type2) +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_KOrS3D2, intercept = intercet_wt_KOrS3D2, 
              linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                 "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT2 H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized KO+S-3D H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th -> pBins_WT_vs_KOrS3D2_Filtered_Targeted
pBins_WT_vs_KOrS3D2_Filtered_Targeted
```

Save to pdf vector.

```{r, save KOrS3D2 vs WT scatterplot vector}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5G_CGI_coloured_WT_vs_KOrS3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_KOrS3D2_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise only the point layer of ggplot

```{r, rasterise KOrS3D2 vs WT scatterplot}
rBins_WT_vs_KOrS3D2_Filtered_Targeted <- rasterize(pBins_WT_vs_KOrS3D2_Filtered_Targeted,
                                                   layers = 'Point', dpi = 400)
```

Show plot as presented in figure.

```{r, display rasterised figure KOrS3D2, fig.width = 2, fig.height = 2}
rBins_WT_vs_KOrS3D2_Filtered_Targeted
```

Save to pdf rasterised scatter plot.

```{r, save KOrS3D2 vs WT scatterplot rasterised}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5G_RASTERIZED_CGI_coloured_WT_vs_KOrS3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_KOrS3D2_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Show the same plot but **without** the log10 scaling of the data.

```{r, fig.width = 2, fig.height = 2}
ggplot(wt_KOrS3D2_bins) +
  aes(x = 10^WT2, y = 10^KOrS3D2, colour = type2) +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_KOrS3D2, intercept = intercet_wt_KOrS3D2, 
              linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 150), ylim = c(0, 150), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                 "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT#2 H3K27me3 signal" ), 
       y = expression("Spike-in normalized KO+S-3D H3K27me3 signal" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th +
  theme(legend.position = c(0.3, 0.92) ) -> pBins_WT_vs_KOrS3D2_Filtered_Targeted_NOLOG10
pBins_WT_vs_KOrS3D2_Filtered_Targeted_NOLOG10
```

Save to pdf vector.

```{r, save KOrS3D2 vs WT scatterplot vector no log}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5G_CGI_NoLog10_WT_vs_KOrS3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_KOrS3D2_Filtered_Targeted_NOLOG10, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise only the point layer of ggplot

```{r, rasterise KOrS3D2 vs WT scatterplot no log10}
rBins_WT_vs_KOrS3D2_NoLog10 <- rasterize(pBins_WT_vs_KOrS3D2_Filtered_Targeted_NOLOG10, 
                                                   layers = 'Point', dpi = 400)
```

Show plot as presented in figure.

```{r, display rasterised figure KOrS3D2 no log10, fig.width = 2, fig.height = 2}
rBins_WT_vs_KOrS3D2_NoLog10
```

Save to pdf rasterised scatter plot.

```{r, save KOrS3D2 vs WT scatterplot rasterised no log10}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5G_RASTERIZED_CGI_NoLog10_WT_vs_KOrS3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_KOrS3D2_NoLog10, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

### Scatter plot KO+SUZ12-L3D vs WT

SUZ12-L 3D mutant H3K27me3 scatter plot.

Get linear fit.
```{r, linear fit WT vs KOrL3D2 }
intercet_wt_KOrL3D2 <- coef( lm(log_bins25kb_K27me3_rescue$KOrL3D2 ~ log_bins25kb_K27me3_rescue$WT2 ) )[1] 
slope_wt_KOrL3D2 <- coef( lm(log_bins25kb_K27me3_rescue$KOrL3D2 ~ log_bins25kb_K27me3_rescue$WT2 ) )[2] 
```

Check linear fit parameters

```{r}
message("WT vs KO+L-3D: intercept = ", round(intercet_wt_KOrL3D2, 3), 
        " slope = ", round(slope_wt_KOrL3D2, 3))
```

Filter the bins like before.
Start by removing the many "rest of the genome" bins

```{r, filter KOrL3D2 ROTG }
indx_dups_ROTG_wt_KOrL3D2 <- which(duplicated(log_bins25kb_K27me3_rescue[log_bins25kb_K27me3_rescue$type2 == "Rest of the genome", c("WT2", "KOrL3D2")]))
row_names_dups_ROTG <- rownames(log_bins25kb_K27me3_rescue[indx_dups_ROTG_wt_KOrL3D2, ]) 
wt_KOrL3D2_bins <- log_bins25kb_K27me3_rescue[!rownames(log_bins25kb_K27me3_rescue) %in% row_names_dups_ROTG,]
```

Filter the "PcG targeted CGI" in KO rescue SUZ12-Long 3D mutant vs WT dataset

```{r, filter KOrL3D2 CGI}
indx_dups_PTC_wt_KOrL3D2 <- which(duplicated(wt_KOrL3D2_bins[wt_KOrL3D2_bins$type2 == "PcG targeted CGI", c("WT2", "KOrL3D2")] ) )
row_names_dups_PTC <- rownames(wt_KOrL3D2_bins[indx_dups_PTC_wt_KOrL3D2, ]) 
wt_KOrL3D2_bins <- wt_KOrL3D2_bins[!rownames(wt_KOrL3D2_bins) %in% row_names_dups_PTC,]
```

And the +1 and -1 flanking bins
```{r, filter KOrL3D2 flanking bins}
indx_dups_P1M1_wt_KOrL3D2 <- which(duplicated(wt_KOrL3D2_bins[wt_KOrL3D2_bins$type2 %in% c("PcG targeted CGI + 1", "PcG targeted CGI + 1"), c("WT2", "KOrL3D2")] ) )
row_names_dups_P1M1 <- rownames(wt_KOrL3D2_bins[indx_dups_P1M1_wt_KOrS3D2, ]) 
wt_KOrL3D2_bins <- wt_KOrL3D2_bins[!rownames(wt_KOrL3D2_bins) %in% row_names_dups_P1M1,]
```

Check number of bins after filtering:

```{r, get KOrL3D2 bins}
num_fltrd_bins_2.5kb <- nrow(wt_KOrL3D2_bins)
message("Number of bins of 2.5kb width to plot: ", num_fltrd_bins_2.5kb)
```

Plot figure S5F

```{r, plot filtered scatter plot KOrL3D2 vs WT, fig.width = 2, fig.height = 2}
ggplot(wt_KOrL3D2_bins) +
  aes(x = WT2, y = KOrL3D2, colour = type2) +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_KOrL3D2, intercept = intercet_wt_KOrL3D2, 
              linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 3.5), ylim = c(0, 3.5), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                 "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT#2 H3K27me3 signal (" ~ log[10] ~ ")" ), 
       y = expression("Spike-in normalized KO+L-3D H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th -> pBins_WT_vs_KOrL3D2_Filtered_Targeted
pBins_WT_vs_KOrL3D2_Filtered_Targeted
```

Save to pdf vector.

```{r, save KOrL3D2 vs WT scatterplot vector}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5F_CGI_coloured_WT_vs_KOrL3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_KOrL3D2_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise only the point layer of ggplot

```{r, rasterise KOrL3D2 vs WT scatterplot}
rBins_WT_vs_KOrL3D2_Filtered_Targeted <- rasterize(pBins_WT_vs_KOrL3D2_Filtered_Targeted,
                                                   layers = 'Point', dpi = 400)
```

Show plot as presented in figure.

```{r, display rasterised figure KOrL3D2, fig.width = 2, fig.height = 2}
rBins_WT_vs_KOrL3D2_Filtered_Targeted
```

Save to pdf rasterised scatter plot.

```{r, save KOrL3D2 vs WT scatterplot rasterised}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5F_RASTERIZED_CGI_coloured_WT_vs_KOrL3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_KOrL3D2_Filtered_Targeted, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Show the same plot but **without** the log10 scaling of the data.

```{r, fig.width = 2, fig.height = 2}
ggplot(wt_KOrL3D2_bins) +
  aes(x = 10^WT2, y = 10^KOrL3D2, colour = type2) +
  geom_point(size = 0.75, alpha = 0.5, stroke = 0) +
  geom_abline(slope = 1, intercept = 0, linewidth = 0.2) +
  geom_abline(slope = slope_wt_KOrS3D2, intercept = intercet_wt_KOrS3D2, 
              linetype = 'dashed', linewidth = 0.2) +
  coord_fixed(ratio = 1, xlim = c(0, 150), ylim = c(0, 150), clip = 'on' ) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0, 0.5) ) ) +
  scale_colour_manual(values = c("PcG targeted CGI" = '#A2234C',
                                 "PcG targeted CGI - 1" = '#203059', 
                                 "PcG targeted CGI + 1" = '#203059',
                                 "Rest of the genome" = '#72817F'), 
                      breaks = c("PcG targeted CGI", 
                                 "PcG targeted CGI + 1",
                                 "Rest of the genome"),
                      labels = c("PcG targeted CGI" = "Targeted", 
                                 "PcG targeted CGI + 1" = "Flanking",
                                 "Rest of the genome" = "Rest of genome") ) +
  labs(x = expression("Spike-in normalized WT#2 H3K27me3 signal" ), 
       y = expression("Spike-in normalized KO+L-3D H3K27me3 signal" ) ) +
  guides(colour = guide_legend(override.aes = list(alpha = 0.99))) +
  scatter_th +
  theme(legend.position = c(0.27, 0.92) ) -> pBins_WT_vs_KOrL3D2_Filtered_Targeted_NOLOG10
pBins_WT_vs_KOrL3D2_Filtered_Targeted_NOLOG10
```

Save to pdf vector.

```{r, save KOrL3D2 vs WT scatterplot vector no log}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5F_CGI_NoLog10_WT_vs_KOrL3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = pBins_WT_vs_KOrL3D2_Filtered_Targeted_NOLOG10, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

Rasterise only the point layer of ggplot

```{r, rasterise KOrL3D2 vs WT scatterplot no log10}
rBins_WT_vs_KOrL3D2_NoLog10 <- rasterize(pBins_WT_vs_KOrL3D2_Filtered_Targeted_NOLOG10, 
                                                   layers = 'Point', dpi = 400)
```

Show plot as presented in figure.

```{r, display rasterised figure KOrL3D2 no log10, fig.width = 2, fig.height = 2}
rBins_WT_vs_KOrL3D2_NoLog10
```

Save to pdf rasterised scatter plot.

```{r, save KOrL3D2 vs WT scatterplot rasterised no log10}
ggsave(path = pdf_dir_fig5,  
       filename = paste0("FigS5F_RASTERIZED_CGI_NoLog10_WT_vs_KOrL3D2_bins_n", num_fltrd_bins_2.5kb, ".pdf"),
       plot = rBins_WT_vs_KOrL3D2_NoLog10, device = cairo_pdf, units = "cm",
       width = 4.2, height = 4.2)
```

## Boxplot H3K27me3 signal in 2.5kb genomic bins

Display the same H3K27me3 signal in 2.5kb genomic bins as boxplots.

Code below is needed to colour the `ggplot2` facet strip text.

```{r, Only colour strips in x-direction}
# requires package ggh4x
CGI_coloured_strip <-
  strip_themed(background_x = elem_list_rect(fill = c(
    '#203059', '#A2234C', '#203059', '#72817F'
  )),
  text_x = elem_list_text(colour = c('white')))

strip_CGI_conversion <- c(
  `PcG targeted CGI - 1` = "Flanking up",
  `PcG targeted CGI` = "Targeted CGI",
  `PcG targeted CGI + 1` = "Flanking down",
  `Rest of the genome` = "Rest of genome"
)
```

Reshape the data for plotting.

```{r}
log_bins25kb_K27me3 |>
  as_tibble() |>
  pivot_longer(cols = c("WT", 'CSex4', 'dex4', 'KO'), names_to = 'Sample', values_to = "log10_H3K27me3") |>
  mutate(Sample = factor(Sample, levels = c("WT", 'CSex4', 'dex4', 'KO') ) ) |>
  mutate(type2 = factor(type2, levels = c('PcG targeted CGI - 1', 'PcG targeted CGI', 'PcG targeted CGI + 1', 'Rest of the genome') ) ) -> endogenous_bins

log_bins25kb_K27me3_rescue |>
  as_tibble() |>
  pivot_longer(cols = c("WT2", 'KOrL4', 'KOrL6', 'KOrS1', 'KOrS3', 'KOrL3D2', 'KOrS3D2'), 
               names_to = 'Sample', values_to = "log10_H3K27me3") |>
  mutate(Sample = factor(Sample, levels = c("WT2", 'KOrL4', 'KOrL6', 'KOrS1', 'KOrS3', 'KOrL3D2', 'KOrS3D2') ) ) |>
  mutate(type2 = factor(type2, 
                        levels = c('PcG targeted CGI - 1', 
                                   'PcG targeted CGI',
                                   'PcG targeted CGI + 1',
                                   'Rest of the genome') )
         ) -> rescue_bins
```

Plot boxplot for panel 5F without the KO rescue lines

```{r, plot endogenous samples, fig.width = 2.75, fig.height = 1.9}
ggplot(endogenous_bins) +
  aes(x = Sample, y = log10_H3K27me3, fill = Sample) +
  facet_wrap2(~type2, nrow = 1, scales = 'fixed', strip = CGI_coloured_strip, 
              labeller = labeller(type2 = strip_CGI_conversion ) ) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.2, show.legend = F) +
  scale_x_discrete(labels = c('dex4' = '∆ex4') ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0.2, 0.2) ) ) +
  coord_cartesian(ylim = c(0, 3.5), clip = 'on') +
  scale_fill_manual(values = c('WT' = "#377AA3", 'CSex4' = "#B65120", 
                               'dex4' = "#F7CB48",'KO' = "#728189") ) +
  labs(y =  expression("Spike-in normalized H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  boxplot_th -> pBoxPlot_CGI_quant_endogenous
pBoxPlot_CGI_quant_endogenous
```

Save to pdf.

```{r, save plot figure 5F endogenous}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5F_2.5kb_Bins_Boxplot_endogenous_lines.pdf"),
       plot = pBoxPlot_CGI_quant_endogenous, device = cairo_pdf, units = "cm",
       width = 7.3, height = 5)
```

Now same plot with rescue lines

```{r, plot rescue samples, fig.width = 3.95, fig.height = 1.9}
ggplot(rescue_bins) +
  aes(x = Sample, y = log10_H3K27me3, fill = Sample) +
  facet_wrap2(~type2, nrow = 1, scales = 'fixed', strip = CGI_coloured_strip, 
              labeller = labeller(type2 = strip_CGI_conversion ) ) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.2, show.legend = F) +
  scale_x_discrete(labels = c('WT2' = 'WT', 'KOrL4' = 'KO+L#1', 
                              'KOrL6' = 'KO+L#2', 'KOrS1' = 'KO+S#1', 
                              'KOrS3' = 'KO+S#2',
                              'KOrL3D2' = 'KO+L-3D', 'KOrS3D2' = 'KO+S-3D') ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02), add = 0),
                     n.breaks = 6) +
  coord_cartesian(ylim = c(0, 2.25), clip = 'on') +
  scale_fill_manual(values = c('WT2' = "#377AA3",
                               "KOrL4" = "mediumpurple2", "KOrL6" = "mediumpurple3", 
                               "KOrS1" = "darkorange1",  "KOrS3" = "darkorange2",
                               "KOrL3D2" = "#74A57F", "KOrS3D2" = "#E8B4BC")
                    ) +
  labs(y =  expression("Spike-in normalized H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  boxplot_th -> pBoxPlot_CGI_quant_rescues
pBoxPlot_CGI_quant_rescues
```

Save to pdf.

```{r, save plot figure 5F rescues}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5F_2.5kb_Bins_Boxplot_rescue_lines.pdf"),
       plot = pBoxPlot_CGI_quant_rescues, device = cairo_pdf, units = "cm",
       width = 7.5, height = 5)
```

Combine the 2 into one plot:

```{r, plot both endogenous and rescue samples, fig.width = 3.95, fig.height = 1.9}
rbind(endogenous_bins, rescue_bins) |>
  subset(! Sample %in% c('KOrL6', 'KOrS3') ) |> #'WT2', 
  ggplot() +
  aes(x = Sample, y = log10_H3K27me3, fill = Sample) +
  facet_wrap2(~type2, nrow = 1, scales = 'fixed', strip = CGI_coloured_strip, 
              labeller = labeller(type2 = strip_CGI_conversion ) ) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.2, show.legend = F) +
  scale_x_discrete(labels = c('dex4' = '∆ex4', 'KOrL4' = 'KO+L', 
                              'KOrS1' = 'KO+S', 'KOrL3D2' = 'KO+L-3D',
                              'KOrS3D2' = 'KO+S-3D') ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0.2, 0.2) ) ) +
  coord_cartesian(ylim = c(0, 3.5), clip = 'on') +
  scale_fill_manual(values = c('WT' = "#377AA3", 'WT2' = "#377AA3", 'CSex4' = "#B65120",
                               'dex4' = "#F7CB48",'KO' = "#728189",
                               "KOrL4" = "mediumpurple2", "KOrS1" = "darkorange1",
                               "KOrL3D2" = "#74A57F", "KOrS3D2" = "#E8B4BC") ) +
  labs(y =  expression("Spike-in normalized H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  boxplot_th -> pBoxPlot_CGI_quant
pBoxPlot_CGI_quant
```

I think the fact that the spike-in normalisation is was done with 2 different batches might explain with the KO rescues are lower than the KO.

```{r, save plot figure 5F}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5F_2.5kb_Bins_Boxplot_endogenous_rescue_lines.pdf"),
       plot = pBoxPlot_CGI_quant, device = cairo_pdf, units = "cm",
       width = 7.5, height = 5)
```

Combine up and down flanking regions into facet

```{r, set facets for combined up and down flanking regions}
CGI_coloured_strip <-
  strip_themed(background_x = elem_list_rect(fill = c(
    '#A2234C', '#203059', '#72817F')),
  text_x = elem_list_text(colour = c('white')))

strip_CGI_conversion <- c(
  `Targeted CGI` = "Targeted CGI",
  `Flanking region` = "Flanking",
  `Rest of the genome` = "Rest of genome"
)
```

Reshape data
```{r, reshape data to split by batch}
samples_batch1 <- c('WT', 'dex4', 'CSex4', 'KO')
samples_batch2 <- c('WT2', 'KOrL4', 'KOrL6', 'KOrS1', 'KOrS3', 'KOrL3D2', 'KOrS3D2')

rbind(endogenous_bins, rescue_bins) |>
  subset(! Sample %in% c('KOrL6', 'KOrS3') ) |>
  mutate(type3 = type2) |>
  mutate(type3 = case_when(type3 == 'PcG targeted CGI - 1' ~ 'Flanking region',
                           type3 == 'PcG targeted CGI + 1' ~ 'Flanking region',
                           type3 == 'PcG targeted CGI' ~ 'Targeted CGI',
                           type3 == 'Rest of the genome' ~ 'Rest of the genome') ) |>
  mutate(type3 = factor(type3, levels = c('Targeted CGI', 'Flanking region', 'Rest of the genome'))) |>
  mutate(batch = case_when(Sample %in% samples_batch1 ~ 'Batch1',
                           Sample %in% samples_batch2 ~ 'Batch2')
         ) -> combined
```

Plot figure 5F final.

```{r, plot both endogenous and rescue samples combined flanking, fig.width = 5.5, fig.height = 1.9}
ggplot(combined) +
  aes(x = Sample, y = log10_H3K27me3, fill = Sample) +
  facet_wrap2(~batch+type3, nrow = 1, scales = 'free_x', strip = CGI_coloured_strip, 
              labeller = labeller(type3 = strip_CGI_conversion ) ) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.2, show.legend = F) +
  scale_x_discrete(labels = c('WT' = 'WT#1', 'WT2' = 'WT#2',
                              'dex4' = '∆ex4', 'KOrL4' = 'KO+L', 
                              'KOrS1' = 'KO+S', 'KOrL3D2' = 'KO+L-3D',
                              'KOrS3D2' = 'KO+S-3D') ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0.2, 0.2) ) ) +
  coord_cartesian(ylim = c(0, 3.5), clip = 'on') +
  scale_fill_manual(values = c('WT' = "#377AA3", 'WT2' = "#377AA3", 'CSex4' = "#B65120",
                               'dex4' = "#F7CB48",'KO' = "#728189",
                               "KOrL4" = "mediumpurple2", "KOrS1" = "darkorange1",
                               "KOrL3D2" = "#74A57F", "KOrS3D2" = "#E8B4BC") ) +
  labs(y =  expression("Spike-in normalized H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  boxplot_th -> pBoxPlot_CGI_quant_3facets
pBoxPlot_CGI_quant_3facets
```

Save to pdf.

```{r, save plot figure 5F 3 facets}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5F_2.5kb_Bins_Boxplot_endogenous_rescue_lines_3facets.pdf"),
       plot = pBoxPlot_CGI_quant_3facets, device = cairo_pdf, units = "cm",
       width = 11.0, height = 5)
```

Split the 2 batches of ChIP-seq in one main and one supplementary figure.

```{r, plot endogenous batch fig 5F, fig.width = 2.5, fig.height = 1.9 }
subset(combined, batch == 'Batch1') |>
ggplot() +
  aes(x = Sample, y = log10_H3K27me3, fill = Sample) +
  facet_wrap2( ~type3, nrow = 1, scales = 'fixed', strip = CGI_coloured_strip, 
              labeller = labeller(type3 = strip_CGI_conversion ) ) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.2, show.legend = F) +
  scale_x_discrete(labels = c('dex4' = '∆ex4') ) +
  scale_y_continuous(expand = expansion(mult = 0, add = c(0.2, 0.2) ) ) +
  coord_cartesian(ylim = c(0, 3.5), clip = 'on') +
  scale_fill_manual(values = c('WT' = "#377AA3", 'CSex4' = "#B65120",
                               'dex4' = "#F7CB48",'KO' = "#728189") ) +
  labs(y =  expression("Spike-in normalized H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  boxplot_th -> pBoxPlot_CGI_quant_3facets_batch1
pBoxPlot_CGI_quant_3facets_batch1
```

Save to pdf.

```{r, save plot figure 5F 3 facets only endogenous batch1 }
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5F_2.5kb_Bins_Boxplot_endogenous_lines_3facets_batch1.pdf"),
       plot = pBoxPlot_CGI_quant_3facets_batch1, device = cairo_pdf, units = "cm",
       width = 5.5, height = 5)
```

Batch 2

```{r, plot rescue batch2 fig 5F, fig.width = 2.85, fig.height = 1.9 }
subset(combined, batch == 'Batch2') |>
ggplot() +
  aes(x = Sample, y = log10_H3K27me3, fill = Sample) +
  facet_wrap2(~type3, nrow = 1, scales = 'fixed', strip = CGI_coloured_strip, 
              labeller = labeller(type3 = strip_CGI_conversion ) ) +
  geom_boxplot(outlier.shape = NA, linewidth = 0.2, show.legend = F) +
  scale_x_discrete(labels = c('WT2' = 'WT', 'KOrL4' = 'KO+L', 'KOrS1' = 'KO+S',
                              'KOrL3D2' = 'KO+L-3D', 'KOrS3D2' = 'KO+S-3D') ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02), add = 0) ) +
  coord_cartesian(ylim = c(0, 2.5), clip = 'on') +
  scale_fill_manual(values = c('WT2' = "#377AA3",
                               "KOrL4" = "mediumpurple2", "KOrS1" = "darkorange1",
                               "KOrL3D2" = "#74A57F", "KOrS3D2" = "#E8B4BC") ) +
  labs(y =  expression("Spike-in normalized H3K27me3 signal (" ~ log[10] ~ ")" ) ) +
  boxplot_th -> pBoxPlot_CGI_quant_3facets_batch2
pBoxPlot_CGI_quant_3facets_batch2
```

```{r, save plot figure 5F 3 facets only rescues batch 2 }
ggsave(path = pdf_dir_fig5, 
       filename = paste0("FigS5H_2.5kb_Bins_Boxplot_rescue_lines_3facets_batch2.pdf"),
       plot = pBoxPlot_CGI_quant_3facets_batch2, device = cairo_pdf, units = "cm",
       width = 5.5, height = 5)
```

## H3K27me3 signal at PRC2 targeted (or not) CpG Islands

CGI = CpG islands. 

Targted = SUZ12 peak was found to overlap the CGI.

Non-targeted = SUZ12 didn't bind that CGI.

### Boxplot spike-in normalised

Import bed files for the boxplots. Here every sample is identified by a number.

```{r, import bed files}
# b <- 1
region <- c()
sample <- c()
beds <- list()
for (b in 1:length(bed_paths)) {
  # define region based on letter
  if ( grepl(pattern = "^[1-4]A", x = bed_paths[b] ) ) { region[b] <- "CGI_H3K27me3_genes" } 
  else if (grepl(pattern = "^[1-4]B", x = bed_paths[b] ) ) { region[b] <- "CGI_nonPcG_genes" }
  # define sample based on number
  if ( grepl(pattern = "^1", x = bed_paths[b]) ) { sample[b] <- "WT" } 
  else if (grepl(pattern = "^2", x = bed_paths[b] ) ) { sample[b] <- "∆ex4" }
  else if (grepl(pattern = "^3", x = bed_paths[b] ) ) { sample[b] <- "KO" }
  else if (grepl(pattern = "^4", x = bed_paths[b] ) ) { sample[b] <- "CSex4" }
  
  message("Importing sample: ", sample[b], ", region: ", region[b])
  # read in each bed file and annotate it
  read_delim(file = file.path(round17_meta_dir, bed_paths[b]), delim = "\t", 
             col_names = c('chr', 'start', 'end', 'min', 'average', 'max', 'sum'),
             show_col_types = FALSE) |>
    mutate(Sample = sample[b], Region = region[b]) |>
    group_by(Sample, Region) |>
    mutate(Numbered_region = n() ) |>
    ungroup() -> beds[[b]]
  
  names(beds)[b] <- paste0(sample[b], "_", region[b])
}
```

Merge list of dataframes (bed files) into one.

```{r, merge bed files into one dataframe}
all_beds <- do.call('rbind', beds) |>
  mutate(Sample = factor(Sample, levels = c("WT", "CSex4", "∆ex4", "KO")) ) |>
  mutate(Pretty_Region = case_when(Region == "CGI_H3K27me3_genes" ~ "PRC2 targeted CGI",
                                   Region == "CGI_nonPcG_genes" ~ "Other CGI") ) |>
  mutate(Pretty_Region = paste0(Pretty_Region, " (n = ", Numbered_region, ")" ) ) |>
  mutate(Pretty_Region = factor(Pretty_Region, 
                                levels = c("PRC2 targeted CGI (n = 2665)", 
                                           "Other CGI (n = 9999)") ) )
```

Set colours for facets strip text 
```{r, set strip colours}
Metaplot_coloured_strip <- strip_themed(background_x = elem_list_rect(fill = c('#A2234C', '#003e1f') ),
                                   text_x = elem_list_text(colour = c('white') ) )

strip_CGI_conversion <- c(
  `PRC2 targeted CGI (n = 2665)` = "SUZ12 Targeted CGI (n = 2665)",
  `Other CGI (n = 9999)` = "Non-targeted CGI (n = 9999)"
)
```

Plot H3K27me3 signal in the CGI targeted regions used for the meta genes later on as a boxplot.

```{r, plot box plot of metagene profiles signal, fig.width = 2.75, fig.height = 1.9}
ggplot(all_beds) +
  aes(x = Sample, y = log(average + 1, base = 10), fill = Sample) +
  facet_wrap2( ~Pretty_Region, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_boxplot(outlier.shape = NA, show.legend = F, linewidth = 0.2) +
  scale_fill_manual(values = c('WT' = "#377AA3", 'CSex4' = "#B65120", 
                               '∆ex4' = "#F7CB48",'KO' = "#728189") ) +
  labs(y = expression("Spike-in norm. average H3K27me3 signal (" ~ log[10] ~ " + 1)" ) ) +
  boxplot_th -> p_metagenes_regions_signal_boxplot
p_metagenes_regions_signal_boxplot
```

Save to pdf.

```{r, save boxplot}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Metagenes_regions_H3K27me3_signal_Boxplot.pdf"),
       plot = p_metagenes_regions_signal_boxplot, device = cairo_pdf, units = "cm",
       width = 7.3, height = 5)
```

### Metagene profiles spike-in normalised in endogenous cells

Import profiles stored in text files.

First import the files Ivano made for the targeted CGI.

```{r, targeted CGI files H3K27me3}
CGI_targeted_K27me3_WT <- read_delim(
  file = file.path(CGI_Dir, "14_MetaCGIplots/7A_GENEplot_5000/GENEprofile_7A_5000.txt"), 
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "PRC2 targeted CGI", Sample = "WT") 

CGI_targeted_K27me3_dex4 <- read_delim(
  file = file.path(CGI_Dir, "14_MetaCGIplots/8A_GENEplot_5000/GENEprofile_8A_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "PRC2 targeted CGI", Sample = "dex4") 

CGI_targeted_K27me3_KO <- read_delim(
  file = file.path(CGI_Dir, "14_MetaCGIplots/9A_GENEplot_5000/GENEprofile_9A_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "PRC2 targeted CGI", Sample = "KO") 

CGI_targeted_K27me3_CSex4 <- read_delim(
  file = file.path(CGI_Dir, "16_MetaCGIplotsEx4100/E3_GENEplot_5000/GENEprofile_E3_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "PRC2 targeted CGI", Sample = "CSex4") 

CGI_targeted_K27me3 <- rbind(CGI_targeted_K27me3_WT, CGI_targeted_K27me3_dex4, CGI_targeted_K27me3_KO, CGI_targeted_K27me3_CSex4) |>
  as_tibble()
```

Then import the non-targeted CGI.

```{r, non targeted CGI files}
CGI_other_K27me3_WT <- read_delim(
  file = file.path(CGI_Dir, "14_MetaCGIplots/7B_GENEplot_5000/GENEprofile_7B_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "Other CGI", Sample = "WT") 

CGI_other_K27me3_dex4 <- read_delim(
  file = file.path(CGI_Dir, "14_MetaCGIplots/8B_GENEplot_5000/GENEprofile_8B_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "Other CGI", Sample = "dex4") 

CGI_other_K27me3_KO <- read_delim(
  file = file.path(CGI_Dir, "14_MetaCGIplots/9B_GENEplot_5000/GENEprofile_9B_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "Other CGI", Sample = "KO") 

CGI_other_K27me3_CSex4 <- read_delim(
  file = file.path(CGI_Dir,"16_MetaCGIplotsEx4100/E1_GENEplot_5000/GENEprofile_E1_5000.txt"),
  delim = '\t', show_col_types = FALSE ) |>
  setNames(c("MetaPosition", "K27me3_signal") ) |>
  mutate(type = "Other CGI", Sample = "CSex4") 

CGI_other_K27me3 <- rbind(CGI_other_K27me3_WT, CGI_other_K27me3_dex4, CGI_other_K27me3_KO, CGI_other_K27me3_CSex4) |>
  as_tibble()
```

Combine data and reshape it for plotting.

```{r, tidy data for meta gene profiles plot}
Meta_K27me3 <- rbind(CGI_targeted_K27me3, CGI_other_K27me3) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Sample, type) |>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Sample = factor(Sample, levels = c('WT', 'CSex4', 'dex4', 'KO') ) ) |>
  mutate(MetaPosition = MetaPosition / 1000)  # turn bases in kilo-bases
```

Rename strip for the figure
```{r, meta gene strip renaming for tarted or not CGI}
strip_CGI_conversion <- c(
  `PRC2 targeted CGI` = "SUZ12 targeted CGI",
  `Other CGI` = "Non-targeted CGI"
)
```

Since the data contains every single nucleotide signal, I use only 1/4 of the data to plot as this does not affect the trend, but makes the pdf file easier to work with. This is done by using ` sample_frac(size = 0.25)`.

```{r, plot figure 5G, fig.width = 2.7, fig.height = 1.85}
Meta_K27me3 |>
  sample_frac(size = 0.25) |> 
ggplot() +
  aes(x = MetaPosition, y = K27me3_signal, colour = Sample) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = 0.025, add = 0 ) ) +
  scale_colour_manual(values = c('WT' = "#377AA3", 'CSex4' = "#B65120",
                                 'dex4' = "#F7CB48",'KO' = "#728189"),
                      labels = c('dex4' = '∆ex4') ) +
  # coord_cartesian(ylim = c(0, NA), clip = 'off' ) +
  labs(x = 'Distance to CGI center (kb)', 
       y = expression("Spike-in normalised H3K27me3 signal (" ~ log[10] ~ ")" )) +
  profile_th -> pMetaPlot_CGI
pMetaPlot_CGI
```

Save to pdf.

```{r, save figure 5F}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5G_Targeted_CGI_Metagene_Profiles_Endogenous_Lines.pdf"),
       plot = pMetaPlot_CGI, device = cairo_pdf, units = "cm",
       width = 7.3, height = 4.9)
```

### Metagene profiles spike-in normalised in rescues cell lines

Metagene profile spike-in normalised.

Import the files Enrique made for the targeted CGI for the MTF2 ChIP-seq.

```{r, import targeted CGI files H3K27me3 in rescues}
trgtd_K27me3_WT1 <- read_prfl(path_dir = round23_metagene_dir,
                              txt_name = "M1A_GENEplot_5000/GENEprofile_M1A_5000.txt",
                              epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "WT1")

trgtd_K27me3_WT2 <- read_prfl(path_dir = round23_metagene_dir,
                              txt_name = "M2A_GENEplot_5000/GENEprofile_M2A_5000.txt",
                              epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "WT2")

trgtd_K27me3_L1 <- read_prfl(path_dir = round23_metagene_dir,
                             txt_name = "M5A_GENEplot_5000/GENEprofile_M5A_5000.txt",
                             epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "KOrL1")

trgtd_K27me3_L2 <- read_prfl(path_dir = round23_metagene_dir,
                             txt_name = "M6A_GENEplot_5000/GENEprofile_M6A_5000.txt",
                             epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "KOrL2")

trgtd_K27me3_S1 <- read_prfl(path_dir = round23_metagene_dir,
                             txt_name = "M3A_GENEplot_5000/GENEprofile_M3A_5000.txt",
                             epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "KOrS1")

trgtd_K27me3_S2 <- read_prfl(path_dir = round23_metagene_dir,
                             txt_name = "M4A_GENEplot_5000/GENEprofile_M4A_5000.txt",
                             epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "KOrS2")

trgtd_K27me3_L3D <- read_prfl(path_dir = round23_metagene_dir,
                             txt_name = "31A_GENEplot_5000/GENEprofile_31A_5000.txt",
                             epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "KOrL3D1")

trgtd_K27me3_S3D <- read_prfl(path_dir = round23_metagene_dir,
                             txt_name = "32A_GENEplot_5000/GENEprofile_32A_5000.txt",
                             epitope = "H3K27me3", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "KOrS3D1")

CGI_targeted_K27me3 <- rbind(trgtd_K27me3_WT1, trgtd_K27me3_WT2,
                             trgtd_K27me3_L1, trgtd_K27me3_L2,
                             trgtd_K27me3_S1, trgtd_K27me3_S2,
                             trgtd_K27me3_L3D, trgtd_K27me3_S3D) |>
  as_tibble()
```

Import rescues signal in not targeted CGIs.

```{r, import non targeted CGI files H3K27me3 in rescues}
nontrgtd_K27me3_WT1 <- read_prfl(path_dir = round23_metagene_dir,
                                txt_name = "M1B_GENEplot_5000/GENEprofile_M1B_5000.txt",
                                  epitope = "H3K27me3", geneset_type = "Other CGI", 
                              sample_name = "WT1")

nontrgtd_K27me3_WT2 <- read_prfl(path_dir = round23_metagene_dir,
                                txt_name = "M2B_GENEplot_5000/GENEprofile_M2B_5000.txt",
                                epitope = "H3K27me3", geneset_type = "Other CGI", 
                                sample_name = "WT2")

nontrgtd_K27me3_L1 <- read_prfl(path_dir = round23_metagene_dir,
                               txt_name = "M5B_GENEplot_5000/GENEprofile_M5B_5000.txt",
                               epitope = "H3K27me3", geneset_type = "Other CGI", 
                               sample_name = "KOrL1")

nontrgtd_K27me3_L2 <- read_prfl(path_dir = round23_metagene_dir,
                               txt_name = "M6B_GENEplot_5000/GENEprofile_M6B_5000.txt",
                               epitope = "H3K27me3", geneset_type = "Other CGI", 
                               sample_name = "KOrL2")

nontrgtd_K27me3_S1 <- read_prfl(path_dir = round23_metagene_dir,
                                txt_name = "M3B_GENEplot_5000/GENEprofile_M3B_5000.txt",
                                epitope = "H3K27me3", geneset_type = "Other CGI", 
                                sample_name = "KOrS1")

nontrgtd_K27me3_S2 <- read_prfl(path_dir = round23_metagene_dir,
                                txt_name = "M4B_GENEplot_5000/GENEprofile_M4B_5000.txt",
                                epitope = "H3K27me3", geneset_type = "Other CGI", 
                                sample_name = "KOrS2")

nontrgtd_K27me3_L3D <- read_prfl(path_dir = round23_metagene_dir,
                                 txt_name = "31B_GENEplot_5000/GENEprofile_31B_5000.txt",
                                 epitope = "H3K27me3", geneset_type = "Other CGI", 
                                 sample_name = "KOrL3D1")

nontrgtd_K27me3_S3D <- read_prfl(path_dir = round23_metagene_dir,
                                 txt_name = "32B_GENEplot_5000/GENEprofile_32B_5000.txt",
                                 epitope = "H3K27me3", geneset_type = "Other CGI", 
                                 sample_name = "KOrS3D1")

CGI_other_K27me3 <- rbind(nontrgtd_K27me3_WT1, nontrgtd_K27me3_WT2,
                          nontrgtd_K27me3_L1, nontrgtd_K27me3_L2,
                          nontrgtd_K27me3_S1, nontrgtd_K27me3_S2,
                          nontrgtd_K27me3_L3D, nontrgtd_K27me3_S3D) |>
  as_tibble()
```

Combine data and reshape it for plotting.

```{r, tidy data of K27me3 for meta gene profiles plot of rescue lines}
Meta_K27me3 <- rbind(CGI_targeted_K27me3, CGI_other_K27me3) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(Genotype = str_remove(pattern = "[a1-z9]$", string = Sample)) |> # remove last number (i.e. replicate)
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Genotype, type, MetaPosition) |>
  mutate(Average_Signal = mean(H3K27me3_signal, na.rm = T) )|>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Genotype = str_replace(pattern = 'r', replacement = '\\+', Genotype)) |>
  mutate(Genotype = factor(Genotype, levels = rev(c('WT', 'KO+L', 'KO+S', 'KO+L3D', 'KO+S3D')) ) ) |>
  mutate(Sample = factor(Sample, 
                         levels = c('WT1', 'WT2', 'KOrL1', 'KOrL2', 'KOrS1', 'KOrS2', 'KOrL3D1', 'KOrS3D1'))) |>
  mutate(MetaPosition = MetaPosition / 1000) |> # turn bases in kilo-bases
# subset a small a fraction of the total.
  select(MetaPosition, type, Genotype, Average_Signal) |> 
  unique() |> sample_frac(size = 0.25) 
```

Set colour palettes and plot

```{r, plot figure S5 H3K27me3 rescues, fig.width = 2.7, fig.height = 1.85}
genotype_palette <- c('WT' = "#377AA3", 'KO+L' = "mediumpurple2",
                      'KO+S' = "darkorange1", 'KO+L3D'  = "#74A57F",
                      'KO+S3D'= "#E8B4BC")

ggplot(Meta_K27me3) +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = 0.1), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.025), add = 0 ) ) +
  scale_colour_manual(values = genotype_palette,
                      guide = guide_legend(reverse = TRUE) ) +
  coord_cartesian(ylim = c(0, NA), clip = 'off' ) +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'Spike-in normalised H3K27me3 signal') +
  profile_th -> pMetaPlot_CGI_H3K27me3_Rescues
pMetaPlot_CGI_H3K27me3_Rescues
```

Save to pdf.

```{r, save figure S5 H3K27me3 rescue metagene profile}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("FigS5I_H3K27me3_Targeted_CGI_Metagene_Profiles_Rescue_Lines.pdf"),
       plot = pMetaPlot_CGI_H3K27me3_Rescues, device = cairo_pdf, units = "cm",
       width = 7.3, height = 4.7)
```


## MTF2 signal at PRC2 targeted (or not) CpG Islands

MTF2 is a component of the PRC2.1 complex.

Metagene profile spike-in normalised.

Import the files Enrique made for the targeted CGI for the MTF2 ChIP-seq.

```{r, import targeted CGI files MTF2}
trgtd_MTF2_WT1 <- read_prfl(path_dir = round21_meta_dir,
                            txt_name = "M1A_GENEplot_5000/GENEprofile_M1A_5000.txt",
                            epitope = "MTF2", geneset_type = "PRC2 targeted CGI", 
                            sample_name = "WT1")

trgtd_MTF2_WT2 <- read_prfl(path_dir = round21_meta_dir,
                            txt_name = "M2A_GENEplot_5000/GENEprofile_M2A_5000.txt",
                            epitope = "MTF2", geneset_type = "PRC2 targeted CGI", 
                            sample_name = "WT2")

trgtd_MTF2_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "M3A_GENEplot_5000/GENEprofile_M3A_5000.txt",
                               epitope = "MTF2", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "CSex4a")

trgtd_MTF2_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "M4A_GENEplot_5000/GENEprofile_M4A_5000.txt",
                               epitope = "MTF2", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "CSex4b")

trgtd_MTF2_dex4a <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "M5A_GENEplot_5000/GENEprofile_M5A_5000.txt",
                              epitope = "MTF2", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "∆ex4a")

trgtd_MTF2_dex4b <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "M6A_GENEplot_5000/GENEprofile_M6A_5000.txt",
                              epitope = "MTF2", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "∆ex4b")

CGI_targeted_MTF2 <- rbind(trgtd_MTF2_WT1, trgtd_MTF2_WT2, 
                           trgtd_MTF2_CSex4a, trgtd_MTF2_CSex4b,
                           trgtd_MTF2_dex4a, trgtd_MTF2_dex4b) |>
  as_tibble()
```

Import MTF2 signal at non-targeted CGI.

```{r, import non targeted CGI files MTF2}
nontrgtd_MTF2_WT1 <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "M1B_GENEplot_5000/GENEprofile_M1B_5000.txt",
                               epitope = "MTF2", geneset_type = "Other CGI", 
                               sample_name = "WT1")

nontrgtd_MTF2_WT2 <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "M2B_GENEplot_5000/GENEprofile_M2B_5000.txt",
                               epitope = "MTF2", geneset_type = "Other CGI", 
                               sample_name = "WT2")

nontrgtd_MTF2_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                  txt_name = "M3B_GENEplot_5000/GENEprofile_M3B_5000.txt",
                                  epitope = "MTF2", geneset_type = "Other CGI", 
                                  sample_name = "CSex4a")

nontrgtd_MTF2_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                  txt_name = "M4B_GENEplot_5000/GENEprofile_M4B_5000.txt",
                                  epitope = "MTF2", geneset_type = "Other CGI", 
                                  sample_name = "CSex4b")

nontrgtd_MTF2_dex4a <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "M5B_GENEplot_5000/GENEprofile_M5B_5000.txt",
                                 epitope = "MTF2", geneset_type = "Other CGI", 
                                 sample_name = "∆ex4a")

nontrgtd_MTF2_dex4b <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "M6B_GENEplot_5000/GENEprofile_M6B_5000.txt",
                                 epitope = "MTF2", geneset_type = "Other CGI", 
                                 sample_name = "∆ex4b")

CGI_other_MTF2 <- rbind(nontrgtd_MTF2_WT1, nontrgtd_MTF2_WT2, 
                        nontrgtd_MTF2_CSex4a, nontrgtd_MTF2_CSex4b,
                        nontrgtd_MTF2_dex4a, nontrgtd_MTF2_dex4b) |>
  as_tibble()
```

Combine data and reshape it for plotting. Since the data contains every single nucleotide signal, I use only 1/4 of the data to plot as this does not affect the trend, but makes the pdf file easier to work with. This is done by using ` sample_frac(size = 0.25)`.

```{r, tidy data of MTF2 for meta gene profiles plot}
Meta_MTF2 <- rbind(CGI_targeted_MTF2, CGI_other_MTF2) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(Genotype = str_remove(pattern = "[a1-z9]$", string = Sample)) |>
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Genotype, type, MetaPosition) |>
  mutate(Average_Signal = mean(MTF2_signal, na.rm = T) )|>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Genotype = factor(Genotype, levels = c('WT', 'CSex4', '∆ex4') ) ) |>
  mutate(Sample = factor(Sample, levels = c('WT1', 'WT2', 'CSex4a', 'CSex4b', '∆ex4a', '∆ex4b') ) ) |>
  mutate(MetaPosition = MetaPosition / 1000) |> # turn bases in kilo-bases
  # subset a small a fraction of the total.
  select(MetaPosition, type, Genotype, Average_Signal) |> 
  unique() |> sample_frac(size = 0.25) 
```

Set colour palettes

```{r, set cell lines labels palette}
samples_palette <- c('WT1' = "#377AA3", 'WT2' = "#377AA3",
                     'CSex4a' = "#B65120", 'CSex4b' = "#B65120",
                     '∆ex4a' = "#F7CB48", '∆ex4b' = "#F7CB48")

genotype_palette <- c('WT' = "#377AA3", 'CSex4' = "#B65120", '∆ex4' = "#F7CB48")
```

Plot MTF2 profile

```{r, plot figure 5 MTF2, fig.width = 2.7, fig.height = 1.85}
ggplot(Meta_MTF2) +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = 0.1), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.025), add = 0 ) ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0, NA), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'Spike-in normalised MTF2 signal') +
  profile_th -> pMetaPlot_CGI_MTF2
pMetaPlot_CGI_MTF2
```

Save to pdf.

```{r, save figure 5 MTF2}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5I_MTF2_Metagene_Profiles.pdf"),
       plot = pMetaPlot_CGI_MTF2, device = cairo_pdf, units = "cm",
       width = 7.3, height = 4.9)
```

Save for main figure.

```{r, save figure 5 MTF2 main panel, fig.width = 2.2, fig.height = 1.85}
Meta_MTF2$Epitope <- 'MTF2'

Meta_MTF2 |>
  subset(type == 'PRC2 targeted CGI') |>
  ggplot() +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~Epitope) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ), n.breaks = 5 ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0, NA), xlim = c(-10, 10), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'Spike-in normalised signal at SUZ12 targeted CGI') +
  profile_th +
  theme(legend.position = c(0.85, 0.87)) -> pMetaPlot_MTF2
pMetaPlot_MTF2
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5I_MTF2_Targeted_CGI_Metagene_Profiles.pdf"),
       plot = pMetaPlot_MTF2, device = cairo_pdf, units = "cm",
       width = 4, height = 4.9)
```

## JARID2 signal at PRC2 targeted (or not) CpG Islands

Metagene profile of spike-in normalised JARID2 ChIP-seq.

```{r, import targeted CGI files JARID2}
trgtd_JARID2_WT1 <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "J1A_GENEplot_5000/GENEprofile_J1A_5000.txt",
                                 epitope = "JARID2", geneset_type = "PRC2 targeted CGI", 
                                 sample_name = "WT1")

trgtd_JARID2_WT2 <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "J2A_GENEplot_5000/GENEprofile_J2A_5000.txt",
                                   epitope = "JARID2", geneset_type = "PRC2 targeted CGI", 
                                   sample_name = "WT2")

trgtd_JARID2_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "J3A_GENEplot_5000/GENEprofile_J3A_5000.txt",
                               epitope = "JARID2", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "CSex4a")

trgtd_JARID2_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "J4A_GENEplot_5000/GENEprofile_J4A_5000.txt",
                               epitope = "JARID2", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "CSex4b")

trgtd_JARID2_dex4a <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "J5A_GENEplot_5000/GENEprofile_J5A_5000.txt",
                              epitope = "JARID2", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "∆ex4a")

trgtd_JARID2_dex4b <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "J6A_GENEplot_5000/GENEprofile_J6A_5000.txt",
                              epitope = "JARID2", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "∆ex4b")

CGI_targeted_JARID2 <- rbind(trgtd_JARID2_WT1, trgtd_JARID2_WT2, 
                             trgtd_JARID2_CSex4a, trgtd_JARID2_CSex4b,
                             trgtd_JARID2_dex4a, trgtd_JARID2_dex4b) |>
  as_tibble()
```

Import JARID2 signal at non-targeted CGI.

```{r, import non targeted CGI files JARID2}
nontrgtd_JARID2_WT1 <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "J1B_GENEplot_5000/GENEprofile_J1B_5000.txt",
                                   epitope = "JARID2", geneset_type = "Other CGI", 
                                   sample_name = "WT1")

nontrgtd_JARID2_WT2 <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "J2B_GENEplot_5000/GENEprofile_J2B_5000.txt",
                                   epitope = "JARID2", geneset_type = "Other CGI", 
                                   sample_name = "WT2")

nontrgtd_JARID2_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "J3B_GENEplot_5000/GENEprofile_J3B_5000.txt",
                                   epitope = "JARID2", geneset_type = "Other CGI", 
                                   sample_name = "CSex4a")

nontrgtd_JARID2_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "J4B_GENEplot_5000/GENEprofile_J4B_5000.txt",
                               epitope = "JARID2", geneset_type = "Other CGI", 
                               sample_name = "CSex4b")

nontrgtd_JARID2_dex4a <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "J5B_GENEplot_5000/GENEprofile_J5B_5000.txt",
                              epitope = "JARID2", geneset_type = "Other CGI", 
                              sample_name = "∆ex4a")

nontrgtd_JARID2_dex4b <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "J6B_GENEplot_5000/GENEprofile_J6B_5000.txt",
                              epitope = "JARID2", geneset_type = "Other CGI", 
                              sample_name = "∆ex4b")

CGI_other_JARID2 <- rbind(nontrgtd_JARID2_WT1, nontrgtd_JARID2_WT2, 
                        nontrgtd_JARID2_CSex4a, nontrgtd_JARID2_CSex4b,
                        nontrgtd_JARID2_dex4a, nontrgtd_JARID2_dex4b) |>
  as_tibble()
```

Combine data and reshape it for plotting.

```{r, tidy data of JARID2 for meta gene profiles plot}
Meta_JARID2 <- rbind(CGI_targeted_JARID2, CGI_other_JARID2) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(Genotype = str_remove(pattern = "[a1-z9]$", string = Sample)) |>
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Genotype, type, MetaPosition) |>
  mutate(Average_Signal = mean(JARID2_signal, na.rm = T) )|>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Genotype = factor(Genotype, levels = c('WT', 'CSex4', '∆ex4') ) ) |>
  mutate(Sample = factor(Sample, levels = c('WT1', 'WT2', 'CSex4a', 'CSex4b', '∆ex4a', '∆ex4b') ) ) |>
  mutate(MetaPosition = MetaPosition / 1000) |> # turn bases in kilo-bases
# subset a small a fraction of the total.
  select(MetaPosition, type, Genotype, Average_Signal) |> 
  unique() |> sample_frac(size = 0.25) 
```

Plot JARID2 profile.

```{r, plot figure 5 JARID2, fig.width = 2.7, fig.height = 1.85}
ggplot(Meta_JARID2) +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.025), add = 0 ) ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0, NA), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'Spike-in normalised JARID2 signal') +
  profile_th -> pMetaPlot_CGI_JARID2
pMetaPlot_CGI_JARID2
```

Save to pdf.

```{r, save figure 5 JARID2}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5H_JARID2_Metagene_Profiles.pdf"),
       plot = pMetaPlot_CGI_JARID2, device = cairo_pdf, units = "cm",
       width = 7.3, height = 4.9)
```

Save for main figure.

```{r, save figure 5 JARID2 main panel, fig.width = 2.2, fig.height = 1.85}
Meta_JARID2$Epitope <- 'JARID2'

Meta_JARID2 |>
  subset(type == 'PRC2 targeted CGI') |>
  ggplot() +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~Epitope) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ), n.breaks = 5 ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0, NA), xlim = c(-10, 10),clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'Spike-in normalised signal at SUZ12 targeted CGI') +
  profile_th +
  theme(legend.position = c(0.85, 0.87)) -> pMetaPlot_JARID2
pMetaPlot_JARID2
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5H_JARID2_Targeted_CGI_Metagene_Profiles.pdf"),
       plot = pMetaPlot_JARID2, device = cairo_pdf, units = "cm",
       width = 4, height = 4.9)
```

## AEBP2 signal at PRC2 targeted (or not) CpG Islands

Import AEBP2 targeted CGI signal not spike-in normalised for metagene profile

```{r, import targeted CGI files AEBP2}
trgtd_AEBP2_WT1 <- read_prfl(path_dir = round21_meta_dir,
                             txt_name = "nA1A_GENEplot_5000/GENEprofile_nA1A_5000.txt",
                             epitope = "AEBP2", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "WT1")

trgtd_AEBP2_WT2 <- read_prfl(path_dir = round21_meta_dir,
                             txt_name = "nA2A_GENEplot_5000/GENEprofile_nA2A_5000.txt",
                             epitope = "AEBP2", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "WT2")

trgtd_AEBP2_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nA3A_GENEplot_5000/GENEprofile_nA3A_5000.txt",
                                epitope = "AEBP2", geneset_type = "PRC2 targeted CGI", 
                                sample_name = "CSex4a")

trgtd_AEBP2_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nA4A_GENEplot_5000/GENEprofile_nA4A_5000.txt",
                                epitope = "AEBP2", geneset_type = "PRC2 targeted CGI", 
                                sample_name = "CSex4b")

trgtd_AEBP2_dex4a <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "nA5A_GENEplot_5000/GENEprofile_nA5A_5000.txt",
                               epitope = "AEBP2", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "∆ex4a")

trgtd_AEBP2_dex4b <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "nA6A_GENEplot_5000/GENEprofile_nA6A_5000.txt",
                               epitope = "AEBP2", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "∆ex4b")

CGI_targeted_AEBP2 <- rbind(trgtd_AEBP2_WT1, trgtd_AEBP2_WT2, 
                            trgtd_AEBP2_CSex4a, trgtd_AEBP2_CSex4b,
                            trgtd_AEBP2_dex4a, trgtd_AEBP2_dex4b) |>
  as_tibble()
```

Import AEBP2 signal at non-targeted CGI.

```{r, import non targeted CGI files AEBP2}
nontrgtd_AEBP2_WT1 <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nA1B_GENEplot_5000/GENEprofile_nA1B_5000.txt",
                                epitope = "AEBP2", geneset_type = "Other CGI", 
                                sample_name = "WT1")

nontrgtd_AEBP2_WT2 <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nA2B_GENEplot_5000/GENEprofile_nA2B_5000.txt",
                                epitope = "AEBP2", geneset_type = "Other CGI", 
                                sample_name = "WT2")

nontrgtd_AEBP2_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "nA3B_GENEplot_5000/GENEprofile_nA3B_5000.txt",
                                   epitope = "AEBP2", geneset_type = "Other CGI", 
                                   sample_name = "CSex4a")

nontrgtd_AEBP2_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "nA4B_GENEplot_5000/GENEprofile_nA4B_5000.txt",
                                   epitope = "AEBP2", geneset_type = "Other CGI", 
                                   sample_name = "CSex4b")

nontrgtd_AEBP2_dex4a <- read_prfl(path_dir = round21_meta_dir,
                                  txt_name = "nA5B_GENEplot_5000/GENEprofile_nA5B_5000.txt",
                                  epitope = "AEBP2", geneset_type = "Other CGI", 
                                  sample_name = "∆ex4a")

nontrgtd_AEBP2_dex4b <- read_prfl(path_dir = round21_meta_dir,
                                  txt_name = "nA6B_GENEplot_5000/GENEprofile_nA6B_5000.txt",
                                  epitope = "AEBP2", geneset_type = "Other CGI", 
                                  sample_name = "∆ex4b")

CGI_other_AEBP2 <- rbind(nontrgtd_AEBP2_WT1, nontrgtd_AEBP2_WT2, 
                         nontrgtd_AEBP2_CSex4a, nontrgtd_AEBP2_CSex4b,
                         nontrgtd_AEBP2_dex4a, nontrgtd_AEBP2_dex4b) |>
  as_tibble()
```

Combine data and reshape it for plotting.

```{r, tidy data of AEBP2 for meta gene profiles plot}
Meta_AEBP2 <- rbind(CGI_targeted_AEBP2, CGI_other_AEBP2) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(Genotype = str_remove(pattern = "[a1-z9]$", string = Sample)) |>
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Genotype, type, MetaPosition) |>
  mutate(Average_Signal = mean(AEBP2_signal, na.rm = T) )|>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Genotype = factor(Genotype, levels = c('WT', 'CSex4', '∆ex4') ) ) |>
  mutate(Sample = factor(Sample, levels = c('WT1', 'WT2', 'CSex4a', 'CSex4b', '∆ex4a', '∆ex4b') ) ) |>
  mutate(MetaPosition = MetaPosition / 1000) |> # turn bases in kilo-bases
  # subset a small a fraction of the total.
  select(MetaPosition, type, Genotype, Average_Signal) |> 
  unique() |> sample_frac(size = 0.25) 
```

Plot AEBP2 profile

```{r, plot figure 5 AEBP2, fig.width = 2.7, fig.height = 1.85}
genotype_palette <- c('WT' = "#377AA3", 'CSex4' = "#B65120", '∆ex4' = "#F7CB48")
ggplot(Meta_AEBP2) +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ), n.breaks = 3 ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0.05, 0.15), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'AEBP2 ChIP-seq signal') +
  profile_th -> pMetaPlot_CGI_AEBP2
pMetaPlot_CGI_AEBP2
```

Save to pdf.

```{r, save figure 5 AEBP2}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("FigS5J_AEBP2_Metagene_Profiles.pdf"),
       plot = pMetaPlot_CGI_AEBP2, device = cairo_pdf, units = "cm",
       width = 6.7, height = 4.7)
```

Plot for supplementary figure

```{r, save figure 5 AEBP2 main panel, fig.width = 2.2, fig.height = 1.85}
Meta_AEBP2$Epitope <- 'AEBP2'

Meta_AEBP2 |>
  subset(type == 'PRC2 targeted CGI') |>
  ggplot() +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~Epitope) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ), n.breaks = 3 ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0.05, 0.15), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'ChIP-seq signal at SUZ12 targeted CGI') +
  profile_th +
  theme(legend.position = c(0.85, 0.87)) -> pMetaPlot_AEBP2
pMetaPlot_AEBP2
ggsave(path = pdf_dir_fig5, 
       filename = paste0("FigS5J_AEBP2_Targeted_CGI_Metagene_Profiles.pdf"),
       plot = pMetaPlot_AEBP2, device = cairo_pdf, units = "cm",
       width = 4, height = 4.9)
```

## RING1B signal at PRC2 targeted (or not) CpG Islands

Import RING1B targeted CGI signal not spike-in normalised for metagene profile

```{r, import targeted CGI files RING1B}
trgtd_RING1B_WT1 <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "nR1A_GENEplot_5000/GENEprofile_nR1A_5000.txt",
                              epitope = "RING1B", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "WT1")

trgtd_RING1B_WT2 <- read_prfl(path_dir = round21_meta_dir,
                              txt_name = "nR2A_GENEplot_5000/GENEprofile_nR2A_5000.txt",
                              epitope = "RING1B", geneset_type = "PRC2 targeted CGI", 
                              sample_name = "WT2")

trgtd_RING1B_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "nR3A_GENEplot_5000/GENEprofile_nR3A_5000.txt",
                                 epitope = "RING1B", geneset_type = "PRC2 targeted CGI", 
                                 sample_name = "CSex4a")

trgtd_RING1B_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "nR4A_GENEplot_5000/GENEprofile_nR4A_5000.txt",
                                 epitope = "RING1B", geneset_type = "PRC2 targeted CGI", 
                                 sample_name = "CSex4b")

trgtd_RING1B_dex4a <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nR5A_GENEplot_5000/GENEprofile_nR5A_5000.txt",
                                epitope = "RING1B", geneset_type = "PRC2 targeted CGI", 
                                sample_name = "∆ex4a")

trgtd_RING1B_dex4b <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nR6A_GENEplot_5000/GENEprofile_nR6A_5000.txt",
                                epitope = "RING1B", geneset_type = "PRC2 targeted CGI", 
                                sample_name = "∆ex4b")

CGI_targeted_RING1B <- rbind(trgtd_RING1B_WT1, trgtd_RING1B_WT2, 
                             trgtd_RING1B_CSex4a, trgtd_RING1B_CSex4b,
                             trgtd_RING1B_dex4a, trgtd_RING1B_dex4b) |>
  as_tibble()
```

Import RING1B signal at non-targeted CGI.

```{r, import non targeted CGI files RING1B}
nontrgtd_RING1B_WT1 <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "nR1B_GENEplot_5000/GENEprofile_nR1B_5000.txt",
                                 epitope = "RING1B", geneset_type = "Other CGI", 
                                 sample_name = "WT1")

nontrgtd_RING1B_WT2 <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "nR2B_GENEplot_5000/GENEprofile_nR2B_5000.txt",
                                 epitope = "RING1B", geneset_type = "Other CGI", 
                                 sample_name = "WT2")

nontrgtd_RING1B_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                    txt_name = "nR3B_GENEplot_5000/GENEprofile_nR3B_5000.txt",
                                    epitope = "RING1B", geneset_type = "Other CGI", 
                                    sample_name = "CSex4a")

nontrgtd_RING1B_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                    txt_name = "nR4B_GENEplot_5000/GENEprofile_nR4B_5000.txt",
                                    epitope = "RING1B", geneset_type = "Other CGI", 
                                    sample_name = "CSex4b")

nontrgtd_RING1B_dex4a <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "nR5B_GENEplot_5000/GENEprofile_nR5B_5000.txt",
                                   epitope = "RING1B", geneset_type = "Other CGI", 
                                   sample_name = "∆ex4a")

nontrgtd_RING1B_dex4b <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "nR6B_GENEplot_5000/GENEprofile_nR6B_5000.txt",
                                   epitope = "RING1B", geneset_type = "Other CGI", 
                                   sample_name = "∆ex4b")

CGI_other_RING1B <- rbind(nontrgtd_RING1B_WT1, nontrgtd_RING1B_WT2, 
                         nontrgtd_RING1B_CSex4a, nontrgtd_RING1B_CSex4b,
                         nontrgtd_RING1B_dex4a, nontrgtd_RING1B_dex4b) |>
  as_tibble()
```

Combine data and reshape it for plotting.

```{r, tidy data of RING1B for meta gene profiles plot}
Meta_RING1B <- rbind(CGI_targeted_RING1B, CGI_other_RING1B) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(Genotype = str_remove(pattern = "[a1-z9]$", string = Sample)) |>
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Genotype, type, MetaPosition) |>
  mutate(Average_Signal = mean(RING1B_signal, na.rm = T) )|>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Genotype = factor(Genotype, levels = c('WT', 'CSex4', '∆ex4') ) ) |>
  mutate(Sample = factor(Sample, levels = c('WT1', 'WT2', 'CSex4a', 'CSex4b', '∆ex4a', '∆ex4b') ) ) |>
  mutate(MetaPosition = MetaPosition / 1000) |> # turn bases in kilo-bases
  # subset a small a fraction of the total.
  select(MetaPosition, type, Genotype, Average_Signal) |> 
  unique() |> sample_frac(size = 0.25) 
```

Plot RING1B profile

```{r, plot figure 5 RING1B, fig.width = 2.7, fig.height = 1.85}
ggplot(Meta_RING1B) +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ) ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0.06, NA), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'RING1B ChIP-seq signal') +
  profile_th -> pMetaPlot_CGI_RING1B
pMetaPlot_CGI_RING1B
```

Save to pdf.

```{r, save figure 5 RING1B}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5J_RING1B_Metagene_Profiles.pdf"),
       plot = pMetaPlot_CGI_RING1B, device = cairo_pdf, units = "cm",
       width = 7.3, height = 4.9)
```

Plot for main figure

```{r, save figure 5 RING1B main panel, fig.width = 2.2, fig.height = 1.85}
Meta_RING1B$Epitope <- 'RING1B'

Meta_RING1B |>
  subset(type == 'PRC2 targeted CGI') |>
  ggplot() +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~Epitope) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ), n.breaks = 6 ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0.05, 0.32), xlim = c(-10, 10), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'ChIP-seq signal at SUZ12 targeted CGI') +
  profile_th +
  theme(legend.position = c(0.85, 0.87)) -> pMetaPlot_RING1B
pMetaPlot_RING1B
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5J_RING1B_Targeted_CGI_Metagene_Profiles.pdf"),
       plot = pMetaPlot_RING1B, device = cairo_pdf, units = "cm",
       width = 4, height = 4.9)
```

## H2Aub signal at PRC2 targeted (or not) CpG Islands

Import H2Aub targeted CGI signal not spike-in normalised for metagene profile

```{r, import targeted CGI files H2Aub}
trgtd_H2Aub_WT1 <- read_prfl(path_dir = round21_meta_dir,
                             txt_name = "nU1A_GENEplot_5000/GENEprofile_nU1A_5000.txt",
                             epitope = "H2Aub", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "WT1")

trgtd_H2Aub_WT2 <- read_prfl(path_dir = round21_meta_dir,
                             txt_name = "nU2A_GENEplot_5000/GENEprofile_nU2A_5000.txt",
                             epitope = "H2Aub", geneset_type = "PRC2 targeted CGI", 
                             sample_name = "WT2")

trgtd_H2Aub_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nU3A_GENEplot_5000/GENEprofile_nU3A_5000.txt",
                                epitope = "H2Aub", geneset_type = "PRC2 targeted CGI", 
                                sample_name = "CSex4a")

trgtd_H2Aub_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                txt_name = "nU4A_GENEplot_5000/GENEprofile_nU4A_5000.txt",
                                epitope = "H2Aub", geneset_type = "PRC2 targeted CGI", 
                                sample_name = "CSex4b")

trgtd_H2Aub_dex4a <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "nU5A_GENEplot_5000/GENEprofile_nU5A_5000.txt",
                               epitope = "H2Aub", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "∆ex4a")

trgtd_H2Aub_dex4b <- read_prfl(path_dir = round21_meta_dir,
                               txt_name = "nU6A_GENEplot_5000/GENEprofile_nU6A_5000.txt",
                               epitope = "H2Aub", geneset_type = "PRC2 targeted CGI", 
                               sample_name = "∆ex4b")

CGI_targeted_H2Aub <- rbind(trgtd_H2Aub_WT1, trgtd_H2Aub_WT2, 
                            trgtd_H2Aub_CSex4a, trgtd_H2Aub_CSex4b,
                            trgtd_H2Aub_dex4a, trgtd_H2Aub_dex4b) |>
  as_tibble()
```

Import H2Aub signal at non-targeted CGI.

```{r, import non targeted CGI files H2Aub}
nontrgtd_H2Aub_WT1 <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "nU1B_GENEplot_5000/GENEprofile_nU1B_5000.txt",
                                 epitope = "H2Aub", geneset_type = "Other CGI", 
                                 sample_name = "WT1")

nontrgtd_H2Aub_WT2 <- read_prfl(path_dir = round21_meta_dir,
                                 txt_name = "nU2B_GENEplot_5000/GENEprofile_nU2B_5000.txt",
                                 epitope = "H2Aub", geneset_type = "Other CGI", 
                                 sample_name = "WT2")

nontrgtd_H2Aub_CSex4a <- read_prfl(path_dir = round21_meta_dir,
                                    txt_name = "nU3B_GENEplot_5000/GENEprofile_nU3B_5000.txt",
                                    epitope = "H2Aub", geneset_type = "Other CGI", 
                                    sample_name = "CSex4a")

nontrgtd_H2Aub_CSex4b <- read_prfl(path_dir = round21_meta_dir,
                                    txt_name = "nU4B_GENEplot_5000/GENEprofile_nU4B_5000.txt",
                                    epitope = "H2Aub", geneset_type = "Other CGI", 
                                    sample_name = "CSex4b")

nontrgtd_H2Aub_dex4a <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "nU5B_GENEplot_5000/GENEprofile_nU5B_5000.txt",
                                   epitope = "H2Aub", geneset_type = "Other CGI", 
                                   sample_name = "∆ex4a")

nontrgtd_H2Aub_dex4b <- read_prfl(path_dir = round21_meta_dir,
                                   txt_name = "nU6B_GENEplot_5000/GENEprofile_nU6B_5000.txt",
                                   epitope = "H2Aub", geneset_type = "Other CGI", 
                                   sample_name = "∆ex4b")

CGI_other_H2Aub <- rbind(nontrgtd_H2Aub_WT1, nontrgtd_H2Aub_WT2, 
                         nontrgtd_H2Aub_CSex4a, nontrgtd_H2Aub_CSex4b,
                         nontrgtd_H2Aub_dex4a, nontrgtd_H2Aub_dex4b) |>
  as_tibble()
```

Combine data and reshape it for plotting.

```{r, tidy data of H2Aub for meta gene profiles plot}
Meta_H2Aub <- rbind(CGI_targeted_H2Aub, CGI_other_H2Aub) |>
  mutate(MetaPosition = MetaPosition - 5000) |>
  mutate(Genotype = str_remove(pattern = "[a1-z9]$", string = Sample)) |>
  mutate(type = factor(type, levels = c("PRC2 targeted CGI", "Other CGI") )) |>
  group_by(Genotype, type, MetaPosition) |>
  mutate(Average_Signal = mean(H2Aub_signal, na.rm = T) )|>
  arrange(MetaPosition) |>
  ungroup() |>
  mutate(Genotype = factor(Genotype, levels = c('WT', 'CSex4', '∆ex4') ) ) |>
  mutate(Sample = factor(Sample, levels = c('WT1', 'WT2', 'CSex4a', 'CSex4b', '∆ex4a', '∆ex4b') ) ) |>
  mutate(MetaPosition = MetaPosition / 1000) |> # turn bases in kilo-bases
  # subset a small a fraction of the total.
  select(MetaPosition, type, Genotype, Average_Signal) |> 
  unique() |> sample_frac(size = 0.25) 
```

Plot H2Aub profile

```{r, plot figure 5 H2Aub, fig.width = 2.7, fig.height = 1.85}
ggplot(Meta_H2Aub) +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~type, strip = Metaplot_coloured_strip,
              labeller = labeller(type = strip_CGI_conversion) ) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ) ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0.025, NA), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'H2AK119ub ChIP-seq signal') +
  profile_th -> pMetaPlot_CGI_H2Aub
pMetaPlot_CGI_H2Aub
```

Save to pdf.

```{r, save figure 5 H2Aub}
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5K_H2AK119ub_Metagene_Profiles.pdf"),
       plot = pMetaPlot_CGI_H2Aub, device = cairo_pdf, units = "cm",
       width = 7.3, height = 4.9)
```

Plot for main figure

```{r, save figure 5 H2AK119ub main panel, fig.width = 2.2, fig.height = 1.85}
Meta_H2Aub$Epitope <- 'H2AK119ub'

Meta_H2Aub |>
  subset(type == 'PRC2 targeted CGI') |>
  ggplot() +
  aes(x = MetaPosition, y = Average_Signal, colour = Genotype) +
  facet_wrap2( ~Epitope) +
  geom_line(linewidth = 0.5) +
  scale_x_continuous(expand = expansion(mult = 0, add = c(0, 0.1) ), 
                     breaks = c(-10, -5, 0, 5, 10) ) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05), add = 0 ), n.breaks = 6 ) +
  scale_colour_manual(values = genotype_palette ) +
  coord_cartesian(ylim = c(0.05, 0.32), clip = 'on') +
  labs(x = 'Distance to CGI center (kb)', 
       y = 'ChIP-seq signal at SUZ12 targeted CGI') +
  profile_th +
  theme(legend.position = c(0.85, 0.87)) -> pMetaPlot_CGI_H2Aub
pMetaPlot_CGI_H2Aub
ggsave(path = pdf_dir_fig5, 
       filename = paste0("Fig5K_H2AK119ub_Targeted_CGI_Metagene_Profiles.pdf"),
       plot = pMetaPlot_CGI_H2Aub, device = cairo_pdf, units = "cm",
       width = 4, height = 4.9)
```

# Supplementary Figure Panels

## Venn diagram SUZ12 peaks

Check the overlap of the genes that contain a peak of SUZ12 in their promoter.

```{r, import gene names of SUZ12 targets}
CSex4 <- read.table(file.path(round3_suz12_peaks_dir, "Suz12_100.txt"))$V1
WT_1 <- read.table(file.path(round3_suz12_peaks_dir, "Suz12_WT1.txt"))$V1
WT_2 <- read.table(file.path(round3_suz12_peaks_dir, "Suz12_WT2.txt"))$V1
dEx4_1 <- read.table(file.path(round3_suz12_peaks_dir, "Suz12_KO1.txt"))$V1
dEx4_2 <- read.table(file.path(round3_suz12_peaks_dir, "Suz12_KO2.txt"))$V1

WT <- WT_1[WT_1%in%WT_2]
dEx4 <- dEx4_1[dEx4_1%in%dEx4_2]

message("Num of genes with SUZ12 peaks:\nWT#1:\t", length(WT_1), '\nWT#2:\t', length(WT_2), 
        '\nWT shared: ', length(WT),
        '\n∆ex4#1:\t', length(dEx4_1), '\n∆ex4#2:\t', length(dEx4_2), '\n∆ex4 shared: ', length(dEx4),
        '\nCSex4:\t', length(CSex4))
```

Save to pdf the peaks overlap as venn diagram

```{r, make venn diagram SUZ12 peaks, message = FALSE, echo = T, eval = FALSE}
draw.venn( WT,CSex4,dEx4,
          title = "SUZ12 target overlap",
          subtitle = "",
          nrtype = "pct",
          # t_f = "Arial",
          xtitle = "WT", #xt_f = "Arial",
          ytitle = "CSex4", #yt_f = "Arial",
          ztitle = "∆ex4", #zt_f = "Arial",
          x_c = "#377AA3", y_c = "#B65120", z_c = "#F7CB48", 
          #nr_f = "Arial", 
          nr_fb = 1, 
          output = 'pdf', width = 500, height = 500, 
          filename = file.path(pdf_dir_fig5, 'FigS5A_SUZ12_targets_VennDiagram.pdf') )
```

Now do the same for H3K27me3

```{r, import gene names of H3K27me3 targets}
CSex4 <- read.table(file.path(round2_h3k27me3_peaks_dir, "H3K27me3_100.txt"))$V1
WT_1 <- read.table(file.path(round2_h3k27me3_peaks_dir, "H3K27me3_WT1.txt"))$V1
WT_2 <- read.table(file.path(round2_h3k27me3_peaks_dir, "H3K27me3_WT2.txt"))$V1
dEx4_1 <- read.table(file.path(round2_h3k27me3_peaks_dir, "H3K27me3_KO1.txt"))$V1
dEx4_2 <- read.table(file.path(round2_h3k27me3_peaks_dir, "H3K27me3_KO2.txt"))$V1

WT <- WT_1[WT_1%in%WT_2]
dEx4 <- dEx4_1[dEx4_1%in%dEx4_2]

message("Num of genes with H3K27me3 peaks:\nWT#1:\t", length(WT_1), '\nWT#2:\t', length(WT_2), 
        '\nWT shared: ', length(WT),
        '\n∆ex4#1:\t', length(dEx4_1), '\n∆ex4#2:\t', length(dEx4_2), '\n∆ex4 shared: ', length(dEx4),
        '\nCSex4:\t', length(CSex4))
```

Plot:
```{r, save to pdf the venn diagram of the H3K27me3, message = FALSE, echo = T, eval = FALSE}
draw.venn(WT,CSex4,dEx4,
          title = "H3K27me3 target overlap",
          subtitle = "", nrtype = "pct",
          xtitle = "WT", #xt_f = "Arial",
          ytitle = "CSex4", #yt_f = "Arial",
          ztitle = "∆ex4", #zt_f = "Arial",
          x_c = "#377AA3", y_c = "#B65120", z_c = "#F7CB48", 
          #nr_f = "Arial", 
          nr_fb = 1, 
          output = 'pdf', width = 500, height = 500, 
          filename = file.path(pdf_dir_fig5, 'FigS5A_H3K27me3_targets_VennDiagram.pdf'))
```

## ChIP-qPCR controls

```{r}
genotype_palette <- c('WT' = "#377AA3", 'CSex4' = "#B65120", '∆ex4' = "#F7CB48", 'KO' = "#728189")
ChIP_qPCR_dir <- file.path(oneDrive_Dir, '11_ChIP/ZChIP_test2')
qPCR_path <- list.files(ChIP_qPCR_dir, pattern = '210603_ZChIP_test2.xlsx', full.names = T)
```

```{r}
WT_samples <- c("WTp", "WT#2")
CSex4_samples <- c("CSex4")
dEx4_samples <- c("∆ex4#2", "∆ex4#3")
KO_samples <- c("KO#1", "KO#2")
```

```{r}
data <- read_excel(qPCR_path, sheet = "melt")
# fix old name
data <- data |>
  mutate(Sample = ifelse(Sample == 'WT4C', yes = 'CSex4', no = Sample)) 
data <- data |>
  mutate(Genotype = case_when(Sample %in% WT_samples ~ 'WT',
                              Sample %in% CSex4_samples ~ 'CSex4',
                              Sample %in% dEx4_samples ~ '∆ex4',
                              Sample %in% KO_samples ~ 'KO') ) |>
  mutate(Genotype = factor(Genotype, levels = c('WT', 'CSex4', '∆ex4','KO') ) ) |>
  mutate(Sample = factor(Sample, levels = unique(data$Sample) ) ) |>
  mutate(Gene = factor(Gene, levels = c("Gsc","Hoxd13","Nes","T (Bra)","Actb","Sox2") ) )
```

Calculate avg for barplot
```{r}
avg <- data %>% 
  group_by(Gene, Sample, Genotype) %>% 
  summarise(value = mean(value), .groups = 'keep')
```

Plot ChIP-qPCR
```{r, fig.width = 2.25, fig.height = 1.85}
ggplot(data)+
  aes(x=Sample, y=value, fill = Genotype) +
  facet_grid(~Gene)+
  geom_col(data = avg, show.legend = F, width = 0.95 )+
  geom_point(size = 1, shape = 21, stroke = 0.2)+
  scale_fill_manual(values = genotype_palette ) +
  scale_x_discrete(labels = c('WT', '', 'CSex4', '', '∆ex4', '', 'KO')) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.02), add = 0)) +
  labs(y = "% of input") +
  theme_classic(base_size = 6, base_family = 'Arial') +
  theme(legend.position = c(0.92, 0.85),
        legend.background = element_blank(),
        legend.title = element_blank(),
        legend.key = element_blank(),
        legend.key.size = unit(1, 'mm'),
        legend.text = element_text(margin = margin(l = -1)),
        panel.grid.major.y = element_line(linewidth = 0.2),
        axis.text = element_text(colour = 'black'),
        axis.text.x = element_text(angle = 45, hjust = 1, margin = margin(t = -1, unit = 'mm')),
        axis.title = element_text(size = 5, colour = 'black'),
        axis.title.x = element_blank(),
        axis.title.y = element_text(margin = margin(r = 0)),
        axis.ticks.length = unit(x = 1, units = 'mm'),
        axis.ticks.x = element_blank(),
        axis.line = element_line(linewidth = 0.2),
        strip.background = element_blank(),
        panel.background = element_blank(),
        plot.background = element_blank()) -> p_ChIP_qCPR
p_ChIP_qCPR
```

```{r}
ggsave(filename = 'FigS5C_ChIP_qPCR_PRC2_targets.pdf', 
       path = pdf_dir_fig5, plot = p_ChIP_qCPR,
       device = cairo_pdf, units = 'mm', width = 60, height = 50)
```

## Individual replicates metagene profiles without spike-in normalisation

## Spie chart distribution H3K27me3 in ∆ex4 ESCs.

Spie chart of the percentage distribution of H3K27me3 peaks in ∆ex4 ESCs. The radial increase corresponds to the fold-change relative to WT ESCs.

Code from Ivano's script

```{r, spie chart, fig.width = 4, fig.height = 4}
# WT vs genome
x <- c(3100,2677,278)
y <- c(4661,3143,1248)
percentX <-x/sum(x)
percentY <-y/sum(y)
percentY <- (percentX/percentY)/sum(percentX/percentY)
names(percentX) <- c("Promoter","Genic","Intergenic")
names(percentY) <- names(percentX)
spie(percentY, percentX, multi = c(0.5, 1, 1.5), lwd = 1,
     pie.labs = TRUE, grid = TRUE, grid.labs = TRUE, scale = TRUE, p1.circle = TRUE,
     bg = c(NA, NA, NA),
     col = c("thistle","palegreen","#728189") )
```

Save to pdf.

```{r, save spie chart}
cairo_pdf(file = file.path(pdf_dir_fig5, 'FigS5E_H3K27me3_peaks_location_SpieChart.pdf'), 
          width = 4, height = 4)
spie(percentY, percentX, multi = c(0.5, 1, 1.5), lwd = 1,
     pie.labs = TRUE, grid = TRUE, grid.labs = TRUE, scale = TRUE, p1.circle = TRUE,
     bg = c(NA, NA, NA),
     col = c("thistle","palegreen","#728189") ) 
dev.off()
```

End of analysis.

# Session Info {#sec-sessioninfo}

```{r, print session info}
sessioninfo::session_info()
```
