---
title: "Quarto markdown analysis template"
author: "Niccolò Arecco"
toc: true
number-sections: true
highlight-style: pygments
format:
  html:
    embed-resources: true
    code-fold: true
    code-copy: true
    theme:
      light: flatly
      dark: darkly
editor: source
bibliography: ../Paper_Analysis_References.bib
---

# Intro

Last code execution: `r format(Sys.time(), '%Y %B %d, %A @ %H:%m:%S')`.

Write a concise introduction.

# Set Up

General info. For references do like this [@Schneider1990].

## Packages

Load packages required for the analysis and suppress any message. Check the @sec-sessioninfo section at the end for package versions.

```{r, Loading required packages, warning = F, message = FALSE}
library(dplyr, warn.conflicts = F, quietly = T)
```

## Functions

Define a plot style.
```{r, label = "ggplot themes", eval = T}

```

Helper functions for the  analysis
```{r, funtions used in this analysis}

```

## Directories & File Paths

Here I organise all the file and directories paths I need to run the analysis and define where to save the processed tables and figures.

```{r, set paths and dicretories, warning = FALSE, message = FALSE}
oneDrive_Dir <- file.path("~/OneDrive - CRG - Centre de Regulacio Genomica/Suz12_AS_project")  
code_dir_fig <- file.path(oneDrive_Dir, "_Code/Fig0")
tbl_dir_fig <- file.path(code_dir_fig, "tables")
pdf_dir_fig <- file.path(code_dir_fig, "pdfs")

if (!dir.exists(pdf_dir_fig)) { dir.create(pdf_dir_fig, recursive = T) }
if (!dir.exists(tbl_dir_fig)) { dir.create(tbl_dir_fig, recursive = T) }
```

# Main Figure Panels

Text and analysis code goes here.

::: {.callout-note}
Some important message 
:::


# Supplementary Figure Panels

End analysis.

# Session Info {#sec-sessioninfo}

```{r, print session info}
sessioninfo::session_info()
```
